---
title: "Greater Silver Smelt in 5a and 14 bootstraps"
author: "Pamela J. Woods"
date: "05/02/2020"
output:
  html_document:
    df_print: paged
---

```{r setup, include==FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, cache = TRUE)
  library(tidyverse)
  library(mfdb)
  mdb<-mfdb('Iceland')
  library(Rgadget)
  #setwd('~/Documents/Hafro/fishvice/19-gss/exploratory_models/gadget/01-new_ass_no_distigfs_mt_s6')
  
    fit<- gadget.fit( gd = '/net/hafkaldi/export/home/haf/pamela/19-gss/exploratory_models/gadget/01-new_ass_no_igfs_mt_s5',
                      recruitment_step_age	= tibble(stock = 'gssimm', age = 5, step = 1), 
                      f.age.range = data.frame(stock = c('gssmat', 'gssimm'), age.min = c(6,6), age.max = c(14,14)),
                      wgts = 'WGTS2')
    
  #load('~/Documents/Hafro/fishvice/19-gss/exploratory_models/gadget/01-new_ass_no_igfs_mt_s6/bootfit_first.Rdata')
  load('/net/hafkaldi/export/home/haf/pamela/19-gss/exploratory_models/gadget/01-new_ass_no_igfs_mt_s5/bootfit.Rdata')
  bootfit.igfs <- bootfit; rm(bootfit)
   
  # bootfit.igfs$params %>%
  #      filter(switc   h=='gss.mat.l50.2019', value > 45) %>% select(model) %>% unlist-> removals
  # 
  # bootfit.igfs$params %>%
  #      filter(switch=='gss.aut.alpha', value > 0.35) %>% select(model) %>% unlist-> removals
  # 
  # tst <- bootfit.igfs %>% purrr::map(function(x){if(is.null(x)){x} else {x %>% filter(model %in% removals[2])}})
  # class(tst) <- c('gadget.fit',class(tst))
  # 
  # tst <- bootfit.igfs %>% purrr::map(function(x){if(is.null(x)){x} else {x %>% filter(model %in% 9)}}) #bad: 43,31, 58, 57, 20, 45, 9, 11, 39,46,19
  # class(tst) <- c('gadget.fit',class(tst))
  
  
  #good reasons to remove bootsraps
  removals <- c(0)
  
  
  bootres.igfs <- 
    bootfit.igfs$res.by.year %>% 
    filter(!(model %in% removals)) %>% 
    group_by(year,model) %>%
    summarise(total.biomass = sum(total.biomass),
              catch = sum(catch),
              #harv.biomass = sum(harv.biomass),
              recruitment = sum(recruitment,na.rm=TRUE)) %>% 
    # left_join(bootfit.igfs$stock.full %>%
    #             filter(length>ref.cm) %>% 
    #             group_by(year,model) %>%
    #             summarise(harv.biomass=sum(number*mean.weight))) %>% 
    group_by(year) %>% 
    summarise(mtb=median(total.biomass),
              utb=quantile(total.biomass,0.95),
              ltb=quantile(total.biomass,0.05),
              uutb=quantile(total.biomass,0.75),
              lltb=quantile(total.biomass,0.25),
              cvtb=sd(total.biomass)/mean(total.biomass),
              mr=median(recruitment,na.rm=TRUE),
              ur=quantile(recruitment,na.rm=TRUE,0.95),
              lr=quantile(recruitment,na.rm=TRUE,0.05),
              uur=quantile(recruitment,na.rm=TRUE,0.75),
              llr=quantile(recruitment,na.rm=TRUE,0.25),
              cvr=sd(recruitment,na.rm=TRUE)/mean(recruitment,na.rm=TRUE),
              # mhb=median(harv.biomass),
              # uhb=quantile(harv.biomass,0.95),
              # lhb=quantile(harv.biomass,0.05),
              # uuhb=quantile(harv.biomass,0.75),
              # llhb=quantile(harv.biomass,0.25),
              # cvhb=sd(harv.biomass)/mean(harv.biomass),
              # mhr=median(catch/harv.biomass),
              # uhr=quantile(catch/harv.biomass,0.95),
              # lhr=quantile(catch/harv.biomass,0.05),
              # uuhr=quantile(catch/harv.biomass,0.75),
              # llhr=quantile(catch/harv.biomass,0.25),
              # cvhr=sd(catch/harv.biomass)/mean(catch/harv.biomass)
              ) %>% 
    left_join(bootfit.igfs$res.by.year %>% 
              filter(!(model %in% removals)) %>% 
                filter(grepl('mat',stock)) %>% 
                group_by(year) %>% 
                summarise(mssb=median(total.biomass),
                          ussb=quantile(total.biomass,0.95),
                          lssb=quantile(total.biomass,0.05),
                          uussb=quantile(total.biomass,0.75),
                          llssb=quantile(total.biomass,0.25),
                          cvssb=sd(total.biomass)/mean(total.biomass),
                          mF=median(F),
                          uF=quantile(F,0.95),
                          lF=quantile(F,0.05),
                          uuF=quantile(F,0.75),
                          llF=quantile(F,0.25),
                          cvF=sd(F)/mean(F))) %>% 
    left_join(fit$res.by.year %>% 
                group_by(year) %>% 
                summarise(ftb = sum(total.biomass),
                          catch = sum(catch),
                          fssb = sum(total.biomass[grepl('mat',stock)]),
                          fr = sum(recruitment,na.rm=TRUE),
                          fF = max(F))) #%>% 
    # left_join(fit$stock.full %>%
    #             filter(length>ref.cm) %>% 
    #             group_by(year) %>%
    #             summarise(fhb=sum(number*mean.weight))) %>%
    #mutate(fhr=catch/fhb)
  
  finres.igfs <- bootres.igfs; rm(bootres.igfs)
  
  
  
  cols <-
    bootfit.igfs$res.by.year %>% 
      filter(!(model %in% removals)) %>% 
    filter(stock== 'gssmat', year == 2019) %>%
    mutate(ssb = total.biomass) %>% 
    select(model, year, ssb) %>% 
    mutate(bssb = fit$res.by.year %>% 
             filter(stock== 'gssmat', year == 2019) %>% 
             select(total.biomass) %>% 
             unlist) %>% 
    mutate(col = ifelse(ssb > bssb, 'red', 'blue')) %>% 
    select(model, col) %>% 
    distinct
  
    
  load('/net/hafkaldi/export/home/haf/pamela/19-gss/exploratory_models/gadget/01-new_ass_no_igfs_mt_s5/retroFit.Rdata')
  
  
  load('/net/hafkaldi/export/home/haf/pamela/19-gss/exploratory_models/gadget/01-new_ass_no_igfs_mt_s5/res.Rdata')
  

```

#### Bootstrap-estimated uncertainty

```{r BSpars, echo=FALSE, fig.width = 9, fig.height= 9, fig.cap = "Greater silver smelt in 5.a and 14. Histogram of parameter estimates from 100 bootstrap samples, the red line indicates the estimate from the base run. Limits on the X axis represent bounds of the parameter search."}

hist.dat <-
  bootfit.igfs$params %>%
  filter(!(model %in% removals))%>%
  bind_rows(bootfit.igfs$params %>%
              filter(model==2) %>% 
              mutate(model=0, 
                     value = lower)) %>% 
  bind_rows(bootfit.igfs$params %>%
              filter(model==2) %>% 
              mutate(model=-1, 
                     value = upper)) %>% 
  filter(optimise==1,
         !grepl('rec\\.[0-9]',switch),
         !grepl('init\\.[0-9]',switch),
         !grepl('scalar',switch)) %>%
   mutate(switch = gsub('^[A-Za-z]+\\.','',switch),
         value = ifelse(grepl('k$|mat1$',switch),0.001*value,value),
         label = forcats::fct_recode(switch,
                                     'beta' = "bbin",
                                     'alpha[g]' = 'gil.alpha',
                                     'l[50][g]' ='gil.l50',
                                     'alpha[l]' = 'lln.alpha',
                                     'l[50][l]' ='lln.l50',
                                     'alpha[t]' = 'bmt.alpha',
                                     'l[50][t]' ='bmt.l50',
                                     'alpha[s]' = 'igfs.alpha',
                                     'l[50][s]'='igfs.l50',
                                     'F[0]'='init.F',
                                     'L[infinity]'='Linf',
                                     'lambda'='mat1',
                                     'l[50]'='mat2',
                                     'sigma[0]'='rec.sd',
                                     'l[0]'='recl'  )) %>% 
  ungroup

   ggplot() + 
     geom_histogram(aes(value), fill='gray', 
                    data = hist.dat %>%
                        mutate(value = ifelse(model == -1, NA, value))
                               )+
     geom_blank(aes(value),data = hist.dat %>%
                        filter(model==-1)) + 
#     geom_histogram(data = hist.dat %>%
#                        filter(model==-1), alpha = 0) + 
     facet_wrap(~label,scale='free',labeller = label_parsed) +
     geom_vline(aes(xintercept=value),
              col='red',
              data=fit$params %>%
                filter(optimise==1,
                       !grepl('rec\\.[0-9]',switch),
                        !grepl('init\\.[0-9]',switch),
                       !grepl('scalar',switch)) %>%
                mutate(switch = gsub('^[A-Za-z]+\\.','',switch),
                       value = ifelse(grepl('k$|mat1$',switch),0.001*value,value),
                       label = forcats::fct_recode(switch,
                                                   'beta' = "bbin",
                                                   'alpha[g]' = 'gil.alpha',
                                                   'l[50][g]' ='gil.l50',
                                                   'alpha[l]' = 'lln.alpha',
                                                   'l[50][l]' ='lln.l50',
                                                   'alpha[t]' = 'bmt.alpha',
                                                   'l[50][t]' ='bmt.l50',
                                                   'alpha[s]' = 'igfs.alpha',
                                                   'l[50][s]'='igfs.l50',
                                                   'F[0]'='init.F',
                                                   'L[infinity]'='Linf',
                                                   'lambda'='mat1',
                                                   'l[50]'='mat2',
                                                   'sigma[0]'='rec.sd',
                                                   'l[0]'='recl' ))) +
   theme_light() +
   labs(x='Parameter value',y='% iterations')
 
 
    ```


```{r BSVB, echo=FALSE, fig.width = 9, fig.height= 9, fig.cap = "Greater silver smelt in 5.a and 14. X-Y scatterplot of the bootstrap estimates of Linfty and k. The red cross indicates the base run estimate."}

bootfit.igfs$params%>%
  filter(!(model %in% removals)) %>%
  filter(grepl('k$|Linf',switch)) %>%
  mutate(switch = gsub('^[A-Za-z]+\\.','',switch),
         value = ifelse(grepl('k$|mat1$',switch),0.001*value,value)) %>%
  select(model,switch,value) %>%
  tidyr::spread(switch, value) %>%
  left_join(cols) %>% 
  ggplot(aes(k,Linf, color = col)) + geom_point() +
  geom_point(col='red',
             data=fit$params %>%
               filter(grepl('k$|Linf',switch)) %>%
               mutate(switch = gsub('^[A-Za-z]+\\.','',switch),
                      value = ifelse(grepl('k$|mat1$',switch),0.001*value,value)) %>%
               select(switch,value) %>%
               tidyr::spread(switch, value),
             pch='+',size=10) +
  theme_light() +
  labs(x='k',y=expression(L[infinity]))

bootfit.igfs$params%>%
  filter(!(model %in% removals)) %>%
  filter(grepl('recl$|rec.sd',switch)) %>%
  mutate(switch = gsub('^[A-Za-z]+\\.','',switch),
         value = ifelse(grepl('k$|mat1$',switch),0.001*value,value)) %>%
  select(model,switch,value) %>%
  tidyr::spread(switch, value) %>%
  left_join(cols) %>% 
  ggplot(aes(recl,rec.sd, color = col)) + geom_point() +
  geom_point(col='red',
             data=fit$params %>%
               filter(grepl('recl$|rec.sd',switch)) %>%
               mutate(switch = gsub('^[A-Za-z]+\\.','',switch),
                      value = ifelse(grepl('k$|mat1$',switch),0.001*value,value)) %>%
               select(switch,value) %>%
               tidyr::spread(switch, value),
             pch='+',size=10) +
  theme_light() +
  labs(x='Rec. length',y='Rec. SD')

```


```{r BSpairs, echo=FALSE, message = FALSE, fig.width = 9, fig.height= 9, fig.cap = "Greater silver smelt in 5a and 14. Pairs-plot of all parameters except those related to the number of recruits and initial number at age."}

bootfit.igfs$params %>%
  filter(!(model %in% removals)) %>% 
  filter(optimise==1,
         !grepl('rec\\.[0-9]',switch),
         !grepl('init\\.[0-9]',switch),
         !grepl('scalar',switch),
         !grepl('mat.l50',switch)) %>%
  mutate(switch = gsub('^[A-Za-z]+\\.','',switch),
         value = ifelse(grepl('k$|mat1$',switch),0.001*value,value),
         label = forcats::fct_recode(switch,
                                     'beta' = "bbin",
                                     'alpha[c]' = 'comm.alpha',
                                     'l[50][c]' ='comm.l50',
                                     'alpha[s]' = 'igfs.alpha',
                                     'l[50][s]'='igfs.l50',
                                     'F[0]'='init.F',
                                     'L[infinity]'='Linf',
                                     'lambda'='mat1',
                                     'l[50]'='mat2',
                                     'sigma[0]'='rec.sd',
                                     'l[0]'='recl' )) %>%
  select(model,switch,value) %>% spread(switch,value) %>% left_join(cols) %>% select(-model) %>%
  GGally::ggpairs(columns = 1:10) +#, ggplot2::aes(colour = col)) +
  theme_light()


```


```{r BSrec, echo=FALSE, fig.width = 9, fig.height= 9, fig.cap = "Greater silver smelt in 5a and 14. Boxplots of annual recruitment (age 5) bootstrap estimates, the red line indicates the estimate from the base run."}

bootfit.igfs$res.by.year %>%
  filter(!(model %in% removals)) %>%
  filter(grepl('imm',stock), year > 1999) %>%
  mutate(recruitment = recruitment/1e6) %>%
  ggplot(aes(year,recruitment,group=round(year))) + geom_boxplot()+#geom_ribbon(aes(ymin=l,ymax=u),fill="gold") + geom_line() +
  geom_line(data=fit$res.by.year %>% filter(stock=='gssimm', year > 1999),aes(y=recruitment/1e6,group=1),col='red') +
  labs(y='Recruitment (in millions)',x='Year') +
  theme_light() + 
  ylim(0,200)


```


```{r BSageparest, echo=FALSE, fig.width = 9, fig.height= 9, fig.cap = "Greater silver smelt in 5a and 14. Boxplots of initial age structure bootstrap estimates, the red line indicates the estimate from the base"}

bootfit.igfs$stock.std%>%
  filter(!(model %in% removals)) %>%
  filter(year == 1990) %>%
  group_by(model,age) %>%
  summarise(number=sum(number)) %>%
  ggplot(aes(age,number/1e6,group=round(age))) + geom_boxplot() +
  geom_line(data=fit$stock.std %>%
              filter(year == 1990)%>%
              group_by(age) %>%
              summarise(number=sum(number)),aes(group=1),col='red') +
  labs(y='Initial age structure (in millions)',x='Age') +
  theme_light()

```


```{r BSqest, echo=FALSE, fig.width = 9, fig.height= 6, fig.cap = "Boxplot of estimated catchability parameters, $q_g$, as a function of the survey index length group."}

bootfit.igfs$sidat %>%
  filter(!(model %in% removals)) %>%
  select(name,intercept) %>%
  #filter(!(name %in% c('si.50-60.aut', 'si.50-60.igfs'))) %>% 
  distinct() %>%
  group_by(name) %>% 
  ggplot(aes(name,exp(intercept))) + 
  geom_boxplot() +
  geom_point() +
  geom_line(aes(group=1),col='red',
            data=fit$sidat %>%
              select(name,intercept) %>%
              distinct()) +
  theme_light() +
  labs(x='Survey index',y=expression(q[g])) #+ 
  #ylim(0,0.0005)
```


```{r BSSI, echo=FALSE, fig.width = 9, fig.height= 6, fig.cap = "Bootstrap distribution of the length aggregated abundance indices from the autumn survey compared with the predicted survey indices. The black line is the bootstrap median and the yellow area is the 2.5 and 97.5% percentiles of the bootstrapped indices, while the black points indicate the survey index.  The blue solid line is the median of the predicted indices from the  bootstrap runs and the blue dotted line the 2.5 and 97.5% percentile.  The red line is the predicted indices from the base model. "}

#fit to individual data sets
#
bootfit.igfs$sidat %>%
#filter(!(model %in% c(22,70,73,24,56,64)))%>%
#filter(name != 'si.45-50.aut') %>% 
group_by(year,name) %>%
  summarise(om=median(observed,na.rm=TRUE),
            ou=quantile(observed,0.975,na.rm=TRUE),
            ol=quantile(observed,0.025,na.rm=TRUE),
            pm=median(predict,na.rm=TRUE),
            pu=quantile(predict,0.975,na.rm=TRUE),
            pl=quantile(predict,0.025,na.rm=TRUE)) %>%
  ggplot(aes(year,om)) + geom_ribbon(aes(ymin=ol,ymax=ou),fill="gold") + geom_line() +
  geom_line(aes(y=pm),col='blue') +
  geom_line(aes(y=pu),col='blue',lty=2) +
  geom_line(aes(y=pl),col='blue',lty=2) +
  geom_line(aes(y=predict),col='red',data=fit$sidat) +
  geom_point(aes(y=observed),data=fit$sidat) +
  facet_wrap(~name,scale='free_y',ncol=2) + theme_light() +
  labs(y='Survey index',x='Year')+
  geom_label(data=fit$sidat %>%
               select(name,slope,sse) %>%
               distinct() %>%
               mutate(label=paste(name,paste0('sse:',sse),sep='\n')),
             aes(label=label),x=-Inf,y=Inf,size=3, vjust = 1.1,hjust=-0.1) +
  theme(strip.background = element_blank(),strip.text=element_blank())

```


```{r BSSIpred, echo=FALSE, fig.width = 9, fig.height= 6, fig.cap = "Predicted survey index as a function of the observed values on the log-scale. The panels indicate the index length group, the dashed line denotes a line with slope 1 and the labels denote the year. "}

fit$sidat %>%
  ggplot(aes(log(predict),log(observed))) +
  geom_line(aes(log(observed),log(observed)),lty=2) +
  geom_line(aes(log(predict),log(predict)),lty=2) +
  facet_wrap(~name, scale='free') +
  geom_text(aes(label=year),size=3) +
  #scale_x_log10() +
  #scale_y_log10() +
  theme_light() +
  labs(y='log(Observed)',x='log(Predicted)')+
  geom_label(data=fit$sidat %>%
               select(name,slope,sse) %>%
               distinct() %>%
               mutate(label=paste(name,paste0('sse:',sse),paste0("slope:",slope),sep='\n')),
             aes(label=label),x=Inf,y=-Inf,size=3, vjust = -0.1,hjust=1.1) +
  theme(strip.background = element_blank(),strip.text=element_blank())


```


```{r BSldist, echo=FALSE, fig.width = 9, fig.height= 6, fig.cap = "Example bootstrap length distribution from both survey and commercial samples compared with model estimates (from 2014, step 4). Green points and vertical bars denote the median and 95% interval of the bootsrap distribution of observed values, while the solid lines and golden ribbon the median and 95% intevals of the bootstrapped estimates by the model. The solid red line indicates the fit from the baseline model."}
bootfit.igfs$catchdist.fleets%>%
  #filter(!(model %in% c(22,70,73,24,56,64))) %>%
  filter(year==2010,step==4) %>%
  mutate(groupy=ifelse(grepl('aldist',name),age,as.character(lower))) %>%
  group_by(name,groupy,model) %>%
  dplyr::summarise(o=sum(observed,na.rm=TRUE),
                   p=sum(predicted,na.rm=TRUE)) %>%
  group_by(name,groupy) %>%
  dplyr::summarise(om=median(o,na.rm=TRUE),
                   pm=median(p,na.rm=TRUE),
                   ou=quantile(o,0.975,na.rm=TRUE),
                   pu=quantile(p,0.975,na.rm=TRUE),
                   ol=quantile(o,0.025,na.rm=TRUE),
                   pl=quantile(p,0.025,na.rm=TRUE)) %>%
  ungroup() %>%
  mutate(groupy=gsub('age','',groupy) %>% as.numeric()) %>%
  ggplot(aes(groupy,om)) +
  geom_ribbon(aes(ymin=pl,ymax=pu),fill='gold') +
  geom_point(col='darkgreen') +
  geom_segment(aes(xend=groupy,yend=ou,y=ol),col='darkgreen') +
  facet_wrap(~name,scale='free') + theme_light() +
  geom_line(aes(y=pm)) + #xlim(c(25,55)) +
  #  geom_label(x=70,y=0.065,aes(label=year),size=3) +
  #  geom_text(x=95,y=0.04,angle=90,aes(label=paste0('n = ',round(total.catch))),size=3) +
  theme(strip.background = element_blank(),strip.text=element_blank()) +
  xlab('Length') + ylab('Proportion') +
  #theme(axis.text.y = element_blank()) +
  geom_label(data=fit$catchdist.fleets %>% select(name) %>% distinct(),
             aes(label=name),x=-Inf,y=Inf,size=3, vjust = 2,hjust=-0.1)


```


```{r BSresnoigfs, echo=FALSE, fig.width = 9, fig.height= 6, fig.cap = "Bootstrap model results wtih NO IGFS INCLUDED. The solid red lines and golden ribbons show the median, 50%, and 90% intervals of the bootstrapped estimates by the model. The solid red black indicates the fit from the baseline model."}



bio.plot <-
  finres.igfs %>% #was bootres
  rename(yea=year) %>%
  select(-matches('cv|r|ssb')) %>%
  rename(year=yea) %>%
  gather(col,value,-year) %>%
  dplyr::mutate(stat=gsub('(^u+|^l+|^m|^f).*',("\\1"),col),
                col = str_sub(col,-2),
                value=value/1e6) %>%
  #filter((col %in% c('hb','tb'))) %>%
  filter((col %in% c('tb'))) %>%
  spread(stat,value) %>%
  dplyr::mutate(col=ifelse(col=='hb','Ref. biomass','Total biomass')) %>%
  ggplot(aes(year,m,group=col, color = col)) + 
  geom_ribbon(aes(ymax=u,ymin=l,fill=col),alpha=0.5) +
  geom_ribbon(aes(ymax=uu,ymin=ll,fill=col),alpha=0.5) +
  geom_line() +
  geom_line(aes(y=f),col='black', linetype = 2)+
  theme_light() +
  expand_limits(y=0) +
  scale_fill_manual(name=NULL,values=c('gold','lightblue')) +
  #scale_color_manual(name=NULL,values=c('gold','lightblue')) +
  labs(y='Biomass (in kt)',x='Year') +
  theme(legend.position = c(0.1,0.9))

ssb.plot <-
  finres.igfs %>%
  select(c(year,matches('ssb'))) %>%
  gather(col,value,-year) %>%
  dplyr::mutate(stat=gsub('(^u+|^l+|^m|^f).*',("\\1"),col),
                col = "ssb",
                value=value/1e6) %>%
  spread(stat,value) %>%
  #filter(year > 1990) %>% 
  ggplot(aes(year,m,group=col, color = col)) +
  geom_ribbon(aes(ymax=u,ymin=l,fill=col),alpha=0.5) +
  geom_ribbon(aes(ymax=uu,ymin=ll,fill=col),alpha=0.5) +
  geom_line() +
  geom_line(aes(y=f),col='black', linetype = 2)+
  theme_light() +
  scale_fill_manual(name=NULL,values=c('gold','lightblue')) +
  expand_limits(y=0) +
  labs(y='SSB (in kt)',x='Year') +
  theme(legend.position = 'none')


F.plot <-
  finres.igfs %>%
#  select(year,matches('[a-z]F|[a-z]hr')) %>%
  select(year,matches('[a-z]F')) %>%
  select(-matches('cv')) %>%
  gather(col,value,-year) %>%
  dplyr::mutate(stat=gsub('(^u+|^l+|^m|^f).*',("\\1"),col),
                col = ifelse(grepl('hr',col),'Harvest rate','Fishing mortality')) %>%
  spread(stat,value) %>%
  #filter(year < thisyear) %>% 
  ggplot(aes(year,m,group=col, color = col)) +
  geom_ribbon(aes(ymax=u,ymin=l,fill=col),alpha=0.5) +
  geom_ribbon(aes(ymax=uu,ymin=ll,fill=col),alpha=0.5) +
  geom_line() +
  geom_line(aes(y=f),col='black', linetype = 2)+
  theme_light() +
  expand_limits(y=0) +
  scale_fill_manual(name=NULL,values=c('gold','lightblue')) +
  #labs(y='Fishing mortality & Harvest rate',x="Year")+
  labs(y='Fishing mortality ',x="Year")+
  theme(legend.position = c(0.15,0.9))

rec.plot <-
  finres.igfs %>%
  select(year,matches('[a-z]r')) %>%
  select(-matches('hr')) %>%
  gather(col,value,-year) %>%
  dplyr::mutate(stat=gsub('(^u+|^l+|^m|^f).*',("\\1"),col),
                col = str_sub(col,-1),
                value = value/1e6) %>%
  filter(stat!='c') %>%
  spread(stat,value) %>%
  ggplot(aes(year,m,group=col, color = col)) +
  geom_ribbon(aes(ymax=u,ymin=l),alpha=0.5,fill='gold') +
  geom_ribbon(aes(ymax=uu,ymin=ll),alpha=0.5,fill='gold') +
  geom_line() +
  geom_line(aes(y=f),col='black', linetype = 2)+
  theme_light() +
  expand_limits(y=0) +
  labs(y='Recruitment age 5 (in millions)',x="Year")

catch.plot <-
  fit$res.by.year %>%
  dplyr::mutate(stock=ifelse(grepl('mat',stock),'Mature','Immature')) %>%
  #filter(year < thisyear) %>% 
  ggplot(aes(year,catch/1e6,fill=stock)) +
  geom_bar(stat='identity') +
  theme_light() +
  scale_fill_manual(name = NULL, values=c('gold','lightblue')) +
  labs(y='Landings (in kt)',x="Year") +
  theme(legend.position = c(0.1,0.9))

cv.plot <-
  finres.igfs %>%
  select(year,matches('cv')) %>%
  gather(col,value,-year) %>%
  dplyr::mutate(stat=str_sub(col,0,2),
         col = str_sub(col,3,5)) %>%
  filter(col %in% c('hb','ssb','r','F'), year > 1999) %>%
  spread(stat,value) %>%
  dplyr::mutate(col=case_when(.$col=='hb'~'Harv. biomass',
                       .$col=='ssb'~'SSB',
                       .$col=='r'~'Recruitment',
                       .$col=='F'~'F')) %>%
  ggplot(aes(year,cv,lty=col)) +
  geom_line() +
  scale_linetype_manual(name=NULL,values=1:4) +
  theme_light() +
  expand_limits(y=0) +
  labs(y='CV',x='Year') +
  theme(legend.position = c(0.2,0.8))

gridExtra::grid.arrange(bio.plot,
                        ssb.plot,# +   
#                          geom_hline(data=data_frame(y=bloss),aes(yintercept=y),col='black',lty=2),
                        F.plot,# +   
#                          geom_hline(data=data_frame(y=c(0.12,0.28)),aes(yintercept=y),col='black',lty=2)+
#                          geom_hline(data=data_frame(y=0.18),aes(yintercept=y),col='black'),
                        rec.plot,
                        catch.plot,ncol=3)

```


```{r retrossi, echo=FALSE, fig.width = 8, fig.height= 8, fig.cap = "Greater silver smelt in 5.a and 14. Analytical retrospective analysis of the fit to the survey indices."}
bootfit.igfs$sidat %>% 
  #filter(!(model %in% c(22,70,73,24,56,64)))%>% 
  group_by(year,name) %>% 
  summarise(om=median(observed,na.rm=TRUE),
            ou=quantile(observed,0.975,na.rm=TRUE),
            ol=quantile(observed,0.025,na.rm=TRUE),
            pm=median(predict,na.rm=TRUE),
            pu=quantile(predict,0.975,na.rm=TRUE),
            pl=quantile(predict,0.025,na.rm=TRUE)) %>% 
  ggplot(aes(year,om)) + geom_ribbon(aes(ymin=ol,ymax=ou),fill="gold") + geom_line() +
  geom_line(aes(y=pm),col='blue') +
  geom_line(aes(y=pu),col='blue',lty=2) +
  geom_line(aes(y=pl),col='blue',lty=2) +
  geom_line(aes(y=predict),col='red',data=fit$sidat) +
  geom_line(aes(y=predict,lty=model),col='red',data=retro.fit$sidat) +
  
  facet_wrap(~name,scale='free_y') + theme_light() + 
  labs(y='Survey index',x='Year')+
  geom_label(data=fit$sidat %>%
              select(name,slope,sse) %>%
              distinct() %>% 
              mutate(label=name),
            aes(label=label),x=-Inf,y=Inf,size=3, vjust = 1.1,hjust=-0.1) +
  theme(strip.background = element_blank(),strip.text=element_blank(),
        legend.position = 'none')

```

### Reference points

According ICES technical guidelines two types of reference points are referred to when giving advice for category 1 stocks: \textit{precautionary approach (PA) reference points and maximum sustainable yield (MSY) reference points. The PA reference points are used when assessing the state of stocks and their exploitation rate relative to the precautionary approach objectives. The MSY reference points are used in the advice rule applied by ICES to give advice consistent with the objective of achieving MSY}.  

The following sections describe the derivation of the management reference points in terms of fishing mortality ($F$). The model for the stock recruitment and assessment error, in combination with the bootstrap results, are used to project the stock status stochastically in order to derive the PA and MSY reference points.  
 
 When considering uncertainty in the estimation of reference points
 four sources of error need to be considered \citep[eg. in][and
 references therein]{butterworth1999}. These are:

 * **Process error** i.e. the natural variation in the stock under
   managment.  Example process that is beyond the control of the
   manager/modeller is variation in recruitment.
 * **Observation error** is relatied to uncertainty in the way the
   stock is sampled.
 * **Implementation error** describes how well the management policies
   are enforced, i.e. the extent of illegal landings and discards.
 * **Model error** related to the form of the model, such as
   stock definitions and variations in model parameter
 
 Observation error and to some degree model error are addressed by the
 bootstrap approach employed in here. 
 Illegal landings and discards by Icelandic fishing vessels are considered to
 be neglible and it is assumed that foreign vessel catches are a part of the 
 management plan. It is assumed therefore implementation error would be virtually none. The largest source of error outstanding is the extent of process
 error, in particular variation in the stock recruitment relationship. 
 
 The following sections describe the model for the stock recruitment and assessment error
 which in combination with the bootstrap results is used to project the
 stock status along with rationale for setting biomass reference points.

#### Recruitment
 The relationship between the estimated spawning stock and recruitment is shown in figure \ref{fig:ssbrec}. The annual recruitment, during the observation period, is estimated
 as a fixed effect without any consideration of the size of the
 spawning stock biomass. This means that no formal stock recruit
 relationship is defined. This is allowed by information on past
 recruitment such as age and length distributions and survey
 indices. It can be observed that for time periods where less data is
 available on recruitment, such as recruitment in the most recent
 years, have higher levels of uncertainty. 
 
 
#### Setting Bloss
The Greater silver smelt stock in Iceland experienced two peaks of heavier fishing pressure: roughly 1998-2000, likely due to a new market opening, and roughly 2009-2011, likely due to a rush to collect fishing history for the species due to its impending entrance into the ITQ system. Therefore, the dip in spawning stock biomass observed roughly 2000 - 2004 is likely to be a result of the first heavy fishing period. Recruitment estimates prior to 2000 are not given as the autumn survey data are only reliable for this species from 2000, and therefore little reliable length distrubution data are available to inform the recruitment estimates prior to this period. 

The stock-recruitment plot for Greater silver smelt shows that there is relatively narrow dynamic range of SSB and no evidence of impaired recruitment, so this stock is categorized as Type 3. In this situation, according to the ICES technical guidelines, $B_{lim}$ can not be estimated from these data and that the lowest observed SSB during that period (i.e. $B_{loss}$ = SSB(2003) = 17.38 kt), which is the lowest observed biomass from the base--line model, and an appropriate value at which to set either as $B_{pa}$ or $B_{lim}$, depending on the the perception of the historical fishing mortality. The fishing mortality is perceived to have since mid 1990s varied between 0.05 and 0.6 on fully recruited, equivalent $F_{6-14}$ of 0.05 to 0.5. As both periods of high fishing mortality are quite short compared to the lifespan of Greater silver smelt. As fishing pressure is perceived to be historically low ,it is suggested that $B_{pa}=B_{loss}$.

In line with ICES technical guidelines, since $B_{pa}$ but not $B_{lim}$ can be estimated, a proxy
for $B_{lim}$ can be calculated based on the inverse of the standard factor, $e^{\sigma * 1.645}$ where $\sigma$ is 0.2, used for calculating $B_{pa}$ from $B_{lim}$. Therefore, a proxy for $B_{lim}$ could be set at $B_{pa}/e^{1.645*0.2} = 17.38/1.39 = 12.50 kt$. 
 

 
```{r SR, echo=FALSE, fig.width = 10, fig.height= 6, fig.cap = "Spawing stock biomass recruitment relationship for Greater silver smelt in 5a adn 14. Uncertainty in recruitment and SSB is indicated with 95% quantile intervals. The yellow vertical bar represents the distribution of $B_{loss}$."}

bloss <- 
  bootfit.igfs$res.by.year %>% 
  filter(grepl('mat',stock)) %>% 
  filter(year > 1999) %>% 
  group_by(model) %>% 
  summarise(bloss=min(total.biomass)/1e6)

tmp <- 
bootfit.igfs$res.by.year %>%
  group_by(year,model) %>% 
  dplyr::summarise(recruitment=sum(recruitment,na.rm=TRUE)/1e6,
                   ssb = total.biomass[grepl('mat',stock)]/1e6) %>% 
  ungroup()

dat <- tmp %>% 
  left_join(tmp %>% 
              select(model,year,ssb.5=ssb) %>% 
              dplyr::mutate(year=year+5)) %>% 
  filter(year > 1999) %>% 
  group_by(year) %>% 
  summarise(mr=median(recruitment),
            ur=quantile(recruitment,0.975),
            lr=quantile(recruitment,0.025),
            ms=median(ssb.5,na.rm=TRUE),
            us=quantile(ssb.5,0.975,na.rm=TRUE),
            ls=quantile(ssb.5,0.025,na.rm=TRUE)) 
dat <- 
  finres.igfs %>% 
  select(year, mr, ur, lr) %>% 
  filter(year > 1999) %>% 
  left_join(finres.igfs %>% 
              mutate(year = year + 5) %>% 
              select(year, ms = mssb, us = ussb, ls = lssb)) %>% 
  mutate(mr = mr/1e6, ur = ur/1e6, lr = lr/1e6, ms = ms/1e6, us = us/1e6, ls = ls/1e6) 
  

dat %>% 
  ggplot(aes(ms,mr)) + 
    
    geom_rect(ymin=-Inf,ymax=Inf,xmin=quantile(bloss$bloss,0.975),
              xmax=quantile(bloss$bloss,0.025),fill='gold')+
    geom_vline(xintercept = median(bloss$bloss),lty=2)+
    geom_vline(xintercept = fit$res.by.year %>% 
                 filter(grepl('mat',stock)) %>%
                 ungroup() %>% 
                 summarise(min(total.biomass)/1e6) %>% 
                 unlist(),col='red', lty = 2)+
    geom_vline(xintercept = 12.50,col='red')+
    geom_errorbar(aes(ymax=ur,ymin=lr),col='grey') + 
    geom_errorbarh(aes(xmax=us,xmin=ls),col='grey') + 
    geom_path()+
    geom_point(col='red')+
    geom_text(aes(label=year-3),vjust=0.15,hjust=0.15) + 
   xlab('Spawning stock biomass (in \'000 tons) ') + 
  ylab('Recruitment (in millions)') + theme_light() +
  expand_limits(x=0,y=0) 
  


```




#### Stock recruitment relationship for projections
A variety of approaches are common when estimating a stock
recruitment relationship. In the absence of a stock-recruitment signal from the available historical data (Fig. \ref{fig:ssbrec}), 
the ICES guidelines suggest that the hockey-stick recruitment function is used (stock Type 3), i.e. 
\begin{equation}
  R_y = \bar{R}_y \min(1,S_y/B_{loss})
\end{equation}
where $R_y$ is annual recruitment, $S_y$ the spawning stock biomass,
$B_{loss}$ the breakpoint in hockey stick function and $\bar{R_y}$ is the
recruitment when not impaired due to low levels of SSB. Here $\bar{R}_y$ is considered to be drawn from the historical distribution using a block-bootstrap, randomly drawn block starting years and including 6 consecutive years in the blocks. This is done to account for intra-correlation in the recruitment time--series, as illustrated in fig. \ref{fig:bootacf}. The timing of the recruitment in the model occurs at the end of the $1^{st}$ time-step, using the projected spawning stock biomass at the same time. 




```{r SRacf, echo=FALSE, fig.width = 10, fig.height= 6, fig.cap = "Histogram of the bootstrap distribution of $B_{loss}$ where the red line indicates the base model estimate."}

bootfit.igfs$res.by.year%>% 
  #filter(!(model %in% removals) %>% 
    filter(grepl('mat',stock)) %>% 
    group_by(model) %>% 
    dplyr::summarise(bloss = min(total.biomass)/1e6) %>% 
    ggplot(aes(bloss)) + geom_histogram() + 
    labs(x=expression(B[loss]),y='Num. replicates') + 
    theme_light() + 
    geom_vline(xintercept = fit$res.by.year %>% 
                 filter(grepl('mat',stock)) %>% 
                 ungroup() %>% 
                 summarise(min=min(total.biomass)/1e6) %>% 
                 unlist(),
               col='red')
```


#### Assessment error 
Observation error and to some degree model error are addressed by the
bootstrap approach employed in here.

Illegal landings and discards by Icelandic fishing vessels are considered to
be negligible and it is assumed that foreign vessel catches will be part of the
management plan when implemented. It is assumed therefore implementation error would be virtually
none. The largest source of error outstanding is the extent of process
error, in particular variation in the stock recruitment relationship, and the assessment error. Derivation of the $B_{pa}$ but not $B_{lim}$ needs no consideration of assessment error, but derivation of $F_{msy}$ and $F_{pa}$ does, so it is introduced here.

The management strategy evaluated was harvesting according to target fishing mortality $F_{target}$. Errors in the assessment
procedure that relate to harvest advice model are emulated as:
\begin{equation}\label{eq:asserr}
\hat{F}_target = e^{E_y} F_{target}
\end{equation}
where $E_y = \sigma (\rho \epsilon_{y-1} + \sqrt{1-\rho^2}\epsilon_y)$ is the assessment error and $\sigma$ is CV of the spawning tock biomass in the most recent estimated year (2019), $\rho$ the autocorrelation between assessment years and $\epsilon_y\sim N(0,1)$. The corresponding CV wasset at UPDATE: 0.25 (i.e, the CV of the  biomass. The autocorrelation in assessment error $\rho$ is set to $0.6$. 

<!-- %\subsection{Residual pattern} -->
<!-- \begin{figure} -->

<!-- <<echo=FALSE,results='asis',fig.width=10, fig.height=6,cache=TRUE,warning=FALSE,message=FALSE>>= -->
<!--   bootacf <-  -->
<!--     bootfit$res.by.year %>%  -->
<!--   #filter(!(model %in% c(22,70,73,24,56,64)))%>% -->
<!--     filter(grepl('imm',stock)) %>%  -->
<!--     group_by(model) %>%  -->
<!--     do(x=acf(.$recruitment,plot=FALSE)) %>%  -->
<!--     broom::tidy(x) %>%  -->
<!--     ggplot(aes(lag,acf,group=round(lag))) + geom_boxplot() +  -->
<!--     theme_light() +  -->
<!--     geom_hline(yintercept = c(-1,1)*qnorm((1 + 0.95)/2)/sqrt(35),lty=2) + -->
<!--     labs(x='',y='') +  -->
<!--     annotate('label',x=13,y=1,label='ACF')  -->

<!--   bootpacf <-  -->
<!--     bootfit$res.by.year %>% -->
<!--     filter(grepl('imm',stock)) %>%  -->
<!--     group_by(model) %>%  -->
<!--     do(x=pacf(.$recruitment,plot=FALSE)) %>%  -->
<!--     broom::tidy(x) %>%  -->
<!--     ggplot(aes(lag,acf,group=round(lag))) + geom_boxplot() +  -->
<!--     theme_light()+  -->
<!--     geom_hline(yintercept = c(-1,1)*qnorm((1 + 0.95)/2)/sqrt(35),lty=2) + -->
<!--     labs(x='',y='') +  -->
<!--     ylim(c(-0.5,1))+  -->
<!--     annotate('label',x=13,y=1,label='PACF')  -->

<!--   bootacf2005 <-  -->
<!--     bootfit$res.by.year %>%  -->
<!--     filter(year < 2005) %>%        -->
<!--     #filter(!(model %in% c(22,70,73,24,56,64)))%>% -->
<!--     filter(grepl('imm',stock)) %>%  -->
<!--     group_by(model) %>%  -->
<!--     do(x=acf(.$recruitment,plot=FALSE)) %>%  -->
<!--     broom::tidy(x) %>%  -->
<!--     ggplot(aes(lag,acf,group=round(lag))) + geom_boxplot() +  -->
<!--     theme_light() +  -->
<!--     geom_hline(yintercept = c(-1,1)*qnorm((1 + 0.95)/2)/sqrt(35),lty=2) + -->
<!--     labs(x='',y='') +  -->
<!--     annotate('label',x=13,y=1,label='ACF')  -->

<!--   bootpacf2005 <-  -->
<!--     bootfit$res.by.year %>% -->
<!--     filter(year < 2005) %>%  -->
<!--     filter(grepl('imm',stock)) %>%  -->
<!--     group_by(model) %>%  -->
<!--     do(x=pacf(.$recruitment,plot=FALSE)) %>%  -->
<!--     broom::tidy(x) %>%  -->
<!--     ggplot(aes(lag,acf,group=round(lag))) + geom_boxplot() +  -->
<!--     theme_light()+  -->
<!--     geom_hline(yintercept = c(-1,1)*qnorm((1 + 0.95)/2)/sqrt(35),lty=2) + -->
<!--     labs(x='',y='') +  -->
<!--     ylim(c(-0.5,1))+  -->
<!--     annotate('label',x=13,y=1,label='PACF')  -->

<!-- gridExtra::grid.arrange(bootacf,bootpacf,bootacf2005,bootpacf2005,ncol=2)   -->
<!-- @ -->

<!-- %\includegraphics[width=1.0\linewidth]{../Graphs/bootacf} -->
<!-- \caption{Estimated autocorrelation (left panels) and partial autocorrelation (right panels) in recruitment. Top panels indicate the correlation including all estimates of recrutitment while the bottom panels recruitment exluding estimates after 2005. The dashed lines illustrate the standard 95\% uncertainty region.}\label{fig:bootacf} -->
<!-- \end{figure} -->

<!-- <<asscomp,echo=FALSE,results='asis',fig.width=10, fig.height=6,cache=TRUE,warning=FALSE,message=FALSE,fig.lp="fig:",fig.cap='ling in 5a. Current assement of the reference biomass of ling (> 75 cm) compared with assessment from the analytical retrospective estimate of the terminal biomass. The shaded yellow ribbon represents the uncertainty (CV = 0.28) at the terminal year.'>>= -->
<!-- fit$res.by.year %>%  -->
<!--   select(year,stock,tru.bio=total.biomass,tru.F=F,tru.rec=recruitment) %>%  -->
<!--   left_join(retro.fit$res.by.year %>%  -->
<!--               filter(year+as.numeric(model)==2016) %>%  -->
<!--               select(year,stock, -->
<!--                      retro.bio = total.biomass, -->
<!--                      retro.F = F, -->
<!--                      retro.rec = recruitment)) %>% -->
<!--   tidyr::gather(type,value,-(year:stock)) %>%  -->
<!--   separate(type,c('source','type')) %>%  -->
<!--   spread(type,value) %>% -->
<!--   group_by(year,source) %>%  -->
<!--   summarise(total.bio=sum(bio)/1e6, -->
<!--             ssb = sum(bio[grepl('mat',stock)])/1e6, -->
<!--             rec = sum(rec,na.rm=TRUE)/1e6) %>%  -->
<!--   left_join(fit$stock.full %>%  -->
<!--               filter(length>ref.cm) %>%  -->
<!--               group_by(year) %>%  -->
<!--               summarise(hb=sum(number*mean.weight)/1e6) %>%  -->
<!--               mutate(source='tru')) %>%  -->
<!--   left_join(retro.fit$stock.full %>%  -->
<!--               filter(length>ref.cm, -->
<!--                      year+as.numeric(model)==2016) %>%  -->
<!--               group_by(year) %>%  -->
<!--               summarise(hb.retro=sum(number*mean.weight)/1e6) %>%  -->
<!--               mutate(source='retro')) %>%  -->
<!--   mutate(hb=ifelse(source=='retro',hb.retro,hb)) %>%  -->
<!--   select(-hb.retro) %>%  -->
<!--   gather(type,value,-c(year:source)) %>%  -->
<!--   filter(type=='hb') %>%  -->
<!--   mutate(value=ifelse(year==2016,value[!is.na(value)&year==2016],value), -->
<!--          source = ifelse(source=='tru','Current assessment','Analytical retrospective')) %>%  -->
<!--   ggplot(aes(year,value,col=source)) +  -->
<!--   geom_ribbon(aes(ymin=value*exp(qnorm(0.025)*0.2), -->
<!--                   ymax=value*exp(qnorm(0.975)*0.2)), -->
<!--               alpha=0.5, -->
<!--               fill='gold',col='white', -->
<!--               data=. %>% filter(source == 'Current assessment'))+ -->
<!--   geom_line() +  -->
<!--   #facet_wrap(~type) + -->
<!--   theme_light() + geom_point() +  -->
<!--   theme(legend.position = c(0.5,0.8)) + -->
<!--   scale_color_manual(name='Assessment',values=c('red','darkblue')) + -->
<!--   labs(y='Reference biomass (in kt)',x ='Year') + -->
<!--   expand_limits(y=0) -->

<!-- @ -->

#### Setting $F_{lim}$ and $F_{pa}$
According to the ICES guidelines, the precautionary reference points and $F_{msy}$ are set by simulating the stock using the stock-recruitment relationship described in section \ref{subsec:SSBR}, based on a wide range of fishing mortality rates, (see eq. \ref{eq:mp}), 
ranging from 0 to 1. For each run of the management strategy evaluation, 1 of the 100 bootstrap runs was forward projected with 1 of 10 randomly generated bootstrapped recruitment, yielding 1000 stochastic simulations. A series of F_{target} values (0 - 1 in steps of 0.01) were then tested by projecting harvesting of each F_{target} for 100 years under in each of the 1000 simulations (100 000 in total).  

<!-- The simulation predicted the stock status was  -->
<!-- projected forward 300 years. For each bootstrap model estimate the stock -->
<!-- status was projected 10 times, resulting in a total of 1000 samples.  -->
<!-- The spawning stock biomass was calculated based on model output after 2060. -->
<!-- This is done to ensure that the stock had reached an equilibrium -->
<!-- under the new fishing mortality regime.  -->

$F_{lim}$ is set as the $F$ that, in equilibrium (last 50 of the simulated 100-year projection), gives a 50\% probability of SSB $> B_{lim}$ without assessment error, described in eq. \ref{eq:asserr}. 
From this $F_{lim}$ is set as the equilibrium fishing mortality when $H_{lim}$ is applied. $H_{pa}$ is then set as the harvest rate that would lead to the equilibrium fishing mortality of $F_{pa}$. $F_{pa}$ is defined as the $F_{lim}/e^{1.645*sigma}$ where $\sigma$ is the CV of the estimated fishing mortality in the assessment year. 


The results from the long term simulations are shown in the top panels of fig. \ref{fig:fmsy}.  $F_{lim}$, resulting in 50% long-term probability of SSB $>B_{lim}$, was estimated at UPDATE. As the CV of $F$ in 2019 is 0.24, $F_{pa}$ is estimated as UPDATE/1.48 = UPDATE. 


#### Maximum sustainable yield
An additional simulation experiment where, in addition to recruitment and bootstrap variations, assessment error was added the harvest rate that would lead to the maximum sustainable yield, $F_{msy}$, was estimated. To calculate $F_{msy}$ ,simmulated annual total landings $c_y$ were calculated over the 50 years from 2070, and plotted over the the F_{target} to form a yield curve. Spawning stock biomasses were similarly analyzed. $F_{msy}$ was then chosen as the $F_{target}$ that corresponds with maximal median yield in the yield curve. 

Figure
\ref{fig:ssbfor} shows the evolution of catches, SSB and fishing mortality for select values of $H$ are shown. The equilibrium yield curve is shown in figure
\ref{fig:fmsy}, where the maximum average yield, under the recruitment
assumptions, is
UPDATE
7.85 thousand tons with a 95\% interval of for the yield 3.64 and
15.00 thousand tons. Table \ref{tbl:prognres} shows the equilibrium results by harvest rate for select statistics.

$F_{msy}$ was estimated to be 0.24.

The
evolution of the spawning stock biomass is shown in figure
\ref{fig:ssbfor}. Equilibrium spawning stock biomass is shown in
figure \ref{fig:fmsy}. The spawning stock biomass obtained at $F_{msy}$ was
estimated at 31.20 thousand tons at $F=0.24$ with an upper quantile
of 58.16 thousand tons and lower quantile of 14.32 thousand tons. 

In line with ICES technical guidelines the MSY $B_{trigger}$ is set as $B_{pa}$, as the stock 
has not been managed according to $F_{msy}$, or equivalents thereof, for more than 5 years. 

```{r yield, echo=FALSE, fig.width = 10, fig.height= 6, fig.cap = "Greater silver smelt in 5a and 14. Equilibrium catch (left) and SSB (right) curves as a function of $F$. Top panels show the results with process error while the bottom panels have furter added assessment uncertainty.  The black solid curves indicate the median projected catch and SSB and the shaded yellow region the 5\% -- 95\% percentiles. Vertical lines indicate $F_{lim}$ (red), $F_{pa}$ (dashed) and $F_{msy}$. The horizontal red line indicates $B_{lim}$."}

bpa <- 
  fit$res.by.year %>% 
  filter(year > 1999) %>% 
  filter(grepl('mat',stock)) %>% 
  ungroup() %>% 
  filter(total.biomass==min(total.biomass)) %>% 
  transmute(total.biomass=total.biomass/1e6) %>% 
  unlist()

blim <- bpa/1.39
bpa <- blim*1.39

Hmsy <- 
  res2 %>% 
  filter(mlnd==max(mlnd)) %>% 
  select(rate,mF)

Hlim <- 
  res20 %>% 
  filter(mssb < blim) %>% 
  head(1) %>% 
  bind_rows(  res20 %>% 
                filter(mssb >= blim) %>% 
                tail(1)) %>%
  mutate(rate=as.numeric(rate)) %>% 
  summarise(rate = rate[2]*(blim-mssb[1])/abs(mssb[1]-mssb[2]) + rate[1]*(mssb[2]-blim)/abs(mssb[1]-mssb[2]),
            mF = mF[2]*(blim-mssb[1])/abs(mssb[1]-mssb[2]) + mF[1]*(mssb[2]-blim)/abs(mssb[1]-mssb[2]))

upper <- Hlim$mF/exp(1.645*0.33)
Hpa <- 
  res20 %>% 
  filter(mF < upper & mF > Hmsy$mF & rate < 0.7) %>% 
  tail(1) %>% 
  bind_rows(  res20 %>% 
                filter(mF >= upper & rate < 0.7) %>% 
                head(1)) %>%
  mutate(rate=as.numeric(rate)) %>% 
  summarise(rate = rate[2]*(upper-mF[1])/abs(mF[1]-mF[2]) + rate[1]*(mF[2]-upper)/abs(mF[1]-mF[2]),
            mF = upper)


yield.plot <- 
  res20 %>% 
  bind_rows(res2,.id='meth') %>% 
  #filter(!(rate %in%c(0.46,0.52))) %>% 
  ggplot(aes(as.numeric(rate),mlnd,group=1)) + 
#  geom_ribbon(aes(ymin=llnd,ymax=ulnd),fill='gold') + 
  geom_ribbon(aes(ymin=mlndq5,ymax=mlndq95),fill='gold',alpha=0.5) + 
  geom_line() +
  #xlim(c(0.01,1)) +
  geom_vline(xintercept = Hmsy$rate %>% as.numeric())+
  geom_vline(xintercept = Hpa$rate, lty=2) +
  geom_vline(xintercept = Hlim$rate, col='red')+
  theme_light() + 
  labs(x='Harvest rate (H)', y='',title='Avg. yield (in kt)') +
  scale_x_continuous(breaks = seq(0,0.5,by=0.1)) + 
  xlim(c(0,0.7)) + 
  facet_wrap(~meth,ncol=1) +
  theme(strip.background = element_blank(),strip.text=element_blank())
ssbmsy <- 
  res20 %>% 
  bind_rows(res2,.id='meth') %>% 
  #filter(!(rate %in%c(0.46,0.52))) %>% 
  ggplot(aes(as.numeric(rate),mssb,group=1)) +
  geom_hline(yintercept =blim,col='red') +
#  geom_ribbon(aes(ymin=lssb,ymax=ussb),fill='gold') + 
  geom_ribbon(aes(ymin=mssbq5,ymax=mssbq95),fill='gold',alpha=0.5) +
  geom_line() +
  #ylim(c(0,)) +
  geom_vline(xintercept = Hmsy$rate %>% as.numeric())+
  geom_vline(xintercept = Hpa$rate, lty=2) +
  geom_vline(xintercept = Hlim$rate, col='red')+
  theme_light() + 
  labs(x='Harvest rate (H)', y='',title='Avg. SSB (in kt)')+
  scale_x_continuous(breaks = seq(0,1,by=0.1)) +
  xlim(c(0,0.7)) + 
  facet_wrap(~meth,ncol=1)  +
  theme(strip.background = element_blank(),strip.text=element_blank())
  
gridExtra::grid.arrange(yield.plot,ssbmsy,ncol=2)
```





```{r exampleFs, eval = F, echo=FALSE, fig.width = 10, fig.height= 6, fig.cap = "Greater silver smelt in 5a and 14. Projected catches, spawning stock biomass and fishing mortality for select target harvest rates. Red and black dashed horizontal lines represent the limit and pa reference points respectively for SSB and F. "}

rates <- c(0.18,0.2,0.22,0.24,0.3)

ssb_progn.plot <- 
res_by_year %>% 
#  bind_rows(.id='rate') %>% 
  filter(rate %in% rates,year < 2050) %>% 
  select(c(year,rate,matches("ssb"))) %>% 
  ggplot(aes(year,mssb/1e6)) + 
  geom_ribbon(aes(ymin=lssb/1e6,ymax=ussb/1e6),fill='gold',alpha=0.5) + 
  geom_ribbon(aes(ymin=llssb/1e6,ymax=uussb/1e6),fill='gold',alpha=0.5) + 
  geom_line() + 
  geom_line(aes(y=ssb6/1e6),lty=2) +
  geom_line(aes(y=ssb542/1e6),lty=3) +
  geom_line(aes(y=ssb931/1e6),lty=4) +
  geom_vline(xintercept = 2017,lty=2) +
  facet_wrap(~rate,ncol=1) +
  theme_light() + 
  labs(title='SSB (in kt)',x='Year',y='') +
  geom_hline(yintercept = bpa,lty=2) +
  geom_hline(yintercept = blim,col='red')+ 
  expand_limits(y=0)

catch_progn.plot <- 
res_by_year %>% 
#  bind_rows(.id='rate') %>% 
  filter(year>2017,rate %in% rates,year < 2050) %>% 
  select(c(year,rate,matches("lnd"))) %>% 
  bind_rows(fit$res.by.year %>% 
              group_by(year) %>% 
              summarise(mlnd=sum(catch/4)) %>% 
              left_join(expand.grid(year=1982:2016,
                                    rate=as.character(rates))))%>% 
  ggplot(aes(year,4*mlnd/1e6)) + 
  geom_ribbon(aes(ymin=4*llnd/1e6,ymax=4*ulnd/1e6),fill='gold',alpha=0.5) + 
  geom_ribbon(aes(ymin=4*lllnd/1e6,ymax=4*uulnd/1e6),fill='gold',alpha=0.5) + 
  geom_line() + 
  geom_line(aes(y=lnd6/1e6),lty=2) +
  geom_line(aes(y=lnd542/1e6),lty=3) +
  geom_line(aes(y=lnd931/1e6),lty=4) +
  geom_vline(xintercept = 2017,lty=2) +
  facet_wrap(~rate,ncol=1) +
  theme_light() + 
  labs(title='Catches (in kt)',x='Year',y='') + 
  expand_limits(y=0)

f_progn.plot <- 
  res_by_year %>% 
  filter(year>2017) %>% 
  bind_rows(bootfit$res.by.year %>% 
              filter(grepl('mat',stock)) %>% 
              group_by(year) %>% 
              dplyr::summarise(mF=mean(F),
                        uuF=quantile(F,0.75),
                        llF=quantile(F,0.25),
                        uF = quantile(F,0.95),
                        lF = quantile(F,0,05)) %>% 
              left_join(expand.grid(year=1982:2016,
                                    rate=as.character(rates))))%>% 
#  bind_rows(.id='rate') %>% 
  filter(rate %in% rates,year < 2050) %>% 
  #select(c(year,rate,)) %>% 
  ggplot(aes(year,mF)) + 
  geom_ribbon(aes(ymin=lF,ymax=uF),fill='gold',alpha=0.5) + 
  geom_ribbon(aes(ymin=llF,ymax=uuF),fill='gold',alpha=0.5) + 
  geom_line() + 
  geom_line(aes(y=F6),lty=2) +
  geom_line(aes(y=F542),lty=3) +
  geom_line(aes(y=F931),lty=4) +
  geom_vline(xintercept = 2017,lty=2) +
  geom_hline(yintercept = Hlim$mF,col='red')+
  geom_hline(yintercept = Hpa$mF,lty=2)+
  facet_wrap(~rate,ncol=1) +
  theme_light() + 
  labs(title='Fishing mortality',x='Year',y='') + 
  expand_limits(y=0)

rec_progn.plot <- 
res_by_year %>% 
  filter(rate %in% rates,year < 2050) %>% 
  select(rate,year,matches('rec')) %>%
  ggplot(aes(year,mrec/1e6)) + 
  geom_ribbon(aes(ymin=lrec/1e6,ymax=urec/1e6),fill='gold',alpha=0.5) + 
  geom_ribbon(aes(ymin=llrec/1e6,ymax=uurec/1e6),fill='gold',alpha=0.5) + 
  geom_line() + 
  geom_line(aes(y=rec6/1e6),lty=2) +
  geom_line(aes(y=rec542/1e6),lty=3) +
  geom_line(aes(y=rec931/1e6),lty=4) +
  geom_vline(xintercept = 2017,lty=2) +
  facet_wrap(~rate,ncol=1) +
  theme_light() + 
  labs(title='Recruitment (in millions)',x='Year',y='') + 
  expand_limits(y=0)

gridExtra::grid.arrange(catch_progn.plot,ssb_progn.plot,f_progn.plot,rec_progn.plot,ncol=4)

```



<!-- <<Fdist,fig.height=8,fig.width=8,fig.lp='fig:',fig.cap="Ling in 5a. Distribution realised harvest rates as a function target harvest rates",echo=FALSE,warning=FALSE,message=FALSE,cache=TRUE>>= -->
<!-- res_quick %>%  -->
<!--   filter(year>2060) %>%  -->
<!--   #filter(rate %in% rates) %>%  -->
<!--   ggplot(aes(hr)) + geom_histogram(aes(y=..density..)) + -->
<!--   theme_light() + #xlim(c(0.1,0.45)) +  -->
<!--   #scale_fill_grey(name='Target harv. rate') + -->
<!--   #theme(legend.position = c(0.9,0.7)) + -->
<!--   labs(x='Realised harvest rate',y='Density') +  -->
<!--   facet_wrap(~rate) + -->
<!--   theme(strip.background = element_blank(), -->
<!--         strip.text=element_blank()) + -->
<!--   geom_label(data=data_frame(rate=rates), -->
<!--              aes(label=rate),  -->
<!--              x=-Inf,y=Inf,size=3,  -->
<!--              vjust = 1.1,hjust=-0.1) + -->
<!--   xlim(c(0,.75)) -->
<!-- @ -->


<!-- \begin{table}[!h] -->
<!-- \vspace{1mm} -->
<!-- \caption[] -->
<!-- {Ling in 5a. Equilibrium averages of yield, fishing mortality (F at age $15^+$) spawning stock biomass (SSB),  recruitment and probility that SSB falls below either $B_{pa}$ or $B_{pa}$ at any point in time from forward projections of the stock status of ling. }\label{tbl:prognres} -->
<!-- \vspace{2mm} -->
<!-- \resizebox{\columnwidth}{!}{\centering -->
<!-- \scriptsize -->
<!-- \begin{tabular}{l| r r r r r r} -->
<!-- \toprule -->
<!-- Harvest rate & Eq. yield & F & SSB & Recruitment & $P(\exists y|SSB_y < B_{pa})$ & $P(\exists y|SSB_y < B_{lim})$ \\ \midrule -->
<!-- <<echo=FALSE,results='asis',message=FALSE,warning=FALSE>>= -->
<!-- res2 %>%  -->
<!--   left_join(res_by_year %>%  -->
<!--               filter(year>2060) %>%  -->
<!--               group_by(rate) %>%  -->
<!--               summarise(mFq5=mean(lF), -->
<!--                         mFq95=mean(uF), -->
<!--                         mrecq5=mean(lrec), -->
<!--                         mrecq95=mean(urec))) %>%  -->
<!--   transmute(harv.rate=as.numeric(rate), -->
<!--             yield = sprintf('%.2f (%.2f,%.2f)',mlnd,mlndq5,mlndq95), -->
<!--             F = sprintf('%.3f (%.3f,%.3f)',mF,mFq5,mFq95), -->
<!--             ssb = sprintf('%.2f (%.2f,%.2f)',mssb,mssbq5,mssbq95), -->
<!--             rec = sprintf('%.2f (%.2f,%.2f)',mrec/1e6,mrecq5/1e6,mrecq95/1e6),#) %>% #, -->
<!--             ppa = sprintf('%.2f',pBpa), -->
<!--             plim = sprintf('%.2f',pBlim)) %>%  -->
<!--   filter(harv.rate < 0.6,harv.rate > 0.05) %>%  -->
<!--   unite(col,matches("."),sep=' & ') %>%  -->
<!--    (function(x){ -->
<!--      paste(x$col,collapse = '\\\\ \n') -->
<!--    }) %>%  -->
<!--    paste0('\\\\') %>%  -->
<!--    cat()  -->
<!-- @ -->

<!-- \bottomrule -->
<!-- \end{tabular} -->
<!-- \par} -->
<!-- \end{table} -->




#### Proposed target harvest rate
ICES guidelines suggest that if $F_{05} < F_{msy}$, where $F_{05}$ is the maximal fishing rate whose median does not exceed a 5% annual risk of the simulated SSB series dropped below $B_{lim}$, then $F_{target}$ should be lowered to $F_{05}$ from $F_{msy}$. As $F_{05}$ was found to be less than $F_{msy}$, it is suggested that $F_{target}$ be reduced to ___. Also, considering the candidate harvest rate for Greater silver smelt in 5a and 14, one may want to investigate the marginal gain of increasing the harvest rate in terms of equilibrium yields. Even with catch rates as low as $F_{05} = $, the equilibrium yields are roughly comparable with that of the $F_{msy}$.  
In addition, short term forward projections illustrated in fig. \ref{fig:ssbfor} indicate that for 
harvest rates will have adapted (on average) to their equilibrium catch levels within a decade from now. 
<!-- Second, the harvest rates have a non-zero chance of exceeding the harvest rate and biomass reference points (as illustrated in fig. \ref{fig:ssbfor} and listed in table \ref{tbl:refpoint}) even at $H_{msy}$, where the probability of SSB going below $B_{lim}$ at any time during the forecast was 0.74, while $H=0.18$ has a probability of 0.05.     -->

The proposed target havest rate, of _ illustrated in fig. \ref{fig:mp.plot}, is considered safe and precautionary as it is lower than any precautionary reference points listed in the sections above. Furthermore it can be considered consistent with the ICES MSY approach, while lower than the estimated $H_{msy}$ of -, the expected long term yield is less that 2% lower. The expected 5% and 95% percentiles of true harvest rates, when setting catches according _, are ... and ... respectively, as illustrated by figure \ref{fig:Fdist}.


UPDATE
Greater silver smelt in 5a and 14. Summary of reference points proposed.. The fishing mortality is relative to ages 6-14. }\label{tbl:refpoint}

Framework   | Reference point | Value | Technical basis
------------|-----------------|-------|-----------------
MSY approach | MSY $B_{trigger}$ |  12.50 | $B_{pa}$
            | $F_{msy}$ | - | The median fishing mortality that maximises the median long-term catch in stochastic simulations with recruitment drawn randomly from ARIMA model prediction uncertainty with AR1 included, scaled according to a hockey stick recruitment function with the breakpoint set to $B_{lim}$ as defined below.
            | $F_{p.05}$ | - |  The median fishing mortality that has an annual probability of 5% of SSB < $B_{lim}$.
------------|-----------------|--------|----------------
Precautionary approach | $B_{lim}$ | 12.50 kt | $B_{pa}/e^{1.645*sigma}$ where $\sigma=0.2$
                      | $B_{pa}$ | 17.38 kt | SSB(2003), corresponding to $B_{loss}$ 
                      | $F_{lim}$ | - | $F$ corresponding to 50% long-term probability of SSB > $B_{lim}$ 
                      | $F_{pa}$ |  - | $F_{lim}/e^{1.645*sigma}$ where $\sigma=0.0.25$
------------|-----------------|--------|----------------
Management plan | $F_{target}$ | - | $F$ such that $F\leq F_{msy}$, long-term yield is consistent with MSY while leading to high stock biomass 
             | $B_{trigger}$ | 12.50 kt | Set as $B_{pa}$ as the stock has not been harvested at $F_{msy}$, or equivalents thereof 
-----------------------------------------------------------

<!-- <<mp.plot, fig.height=5,fig.width=10,fig.lp='fig:',fig.cap='Ling in 5a. Graphical presention of the proposed managment rule. The black solid line indicates the harvest rate relative to the >75cm biomass as a function of the SSB',echo=FALSE>>= -->
<!-- Hmp <- 0.18 -->
<!-- ggplot(aes(x,y), -->
<!--        data=data_frame(x=c(0,bpa,bpa*5),y=c(0,Hmp,Hmp))) +  -->
<!--   geom_path() +  -->
<!--   expand_limits(y=Hmp*1.5) + -->
<!--   theme_light() +  -->
<!--   geom_vline(xintercept = bpa,lty=2) + -->
<!--   geom_vline(xintercept = blim,col='red') + -->
<!--   annotate('text',label=sprintf('Btrigger = %.2f',bpa),x=bpa*1.1,y=0.05,angle=90) +  -->
<!--   labs(y='Harv. rate',x='SSB (in kt)')+ -->
<!--   annotate('text',label=sprintf('Blim = %.2f',blim),x=blim*0.9,y=0.05,angle=90) + -->
<!--   annotate('text',label=sprintf('Hmp = %.2f',Hmp),x=bpa*2,y=1.05*Hmp) -->
<!-- @ -->

### Application of reference points in management

The decision which allocates catches to the fleets requires 1) an expected quanitity of catch to be removed that will complete total catch removals for the current fishing season (not implemented currently, but will be for assessment purposes, 2) a 1-year projection to determine the amount of biomass avaible to fish, and 3) application of projected fishing effort according to $F_{target}$ to determine the expected catch from fishing at this level. The expected catch is then obtained by rearranging the fishing mortality equation and summing over all ages $a$, stocks $s$ and time steps $t$ to form the annual advised TAC:

$$
F_{target, asyt} = \frac{-\log(1.0 - \frac{C_{asyt}}{N_{asyt}})}{\Delta t}
$$
$$
TAC_y = \sum_{ast} C_{y} = \sum_{ast} (exp(F_{target, asyt}\Delta t) +1) N_{asyt}  = C_{asyt}
$$
while $SSB_y > B_{pa}$,  
$$
TAC_y = C_{asyt}*\frac{SSB_y}{B_{trigger}}
$$
while $B_{trigger} \leqslant SSB_y < B_{lim}$, and 
$$
TAC_y = 0
$$
while $B_{lim} \leqslant SSB_y$

### Conclusions
The gadget model presented here captures the overall trends in
the data, and in spite of minor misfits the model is usable for
assessing the stock and to base advice to managers.

In a complicated such as the gadget model that has many parameters and
many data-sets of varying quality it is to be expected that there may
be problems with some parameters and fit to some data sets.

The main problem the model has, as would any other models,
is the rapid increase in the survey indices in recent years, recent uncertain recruitment spikes,  and the
large CV of these indices in that period; however, this pattern may resolve itself as more information on length distributions is gathered

Most parameters are well defined except for the beta-binomial parameter that affects the degree of variation in growth implemented from one time-step to the next. This parameter is often not well defined in Gadget models because there are multiple sources of variation control variability in growth, including standard deviation in recruitment length and the fixed parameter in the beta-binomial distribution that controls the number of maximum length group steps that can be taken in a single time step.

A number of growth parameters are highly correlated in the bootstraps, including $L_{inf}$ and $k$ from the Von Bertalanffy growth equation and the length and standard deviations at recruitment $l_0$ and $sigma_0$. Although correlations in growth parameters are not unexpected, resulting growth patterns should be scrutinized annually to ensure that these constraints do not add bias to the model, as growth is highly influential in population dynamics. Under current model settings and data sources, they predicted mean growth and variation fit patterns in the data quite well, so they are not considered a problem, but should be evaluated in the next benchmark.

The relationship between the estimated spawning stock and recruitment
is shown in figure \ref{fig:ssbrec}. In the Gadget model for Greater silver smelt in
5a and 14, the annual recruitment, during the observation period, is estimated
as a fixed effect without any consideration of the size of the
spawning stock biomass. This means that no formal stock recruit
relationship is defined. This is allowed by information on past
recruitment such as age and length distributions and survey
indices. It can be observed that for time periods where less data is
available on recruitment, such as recruitment in the most recent
years, have higher levels of uncertainty. According to the model, Greater silver smelt in 5a and 14 has experienced increased recruitment in recent years. This pattern could be due to migration or increased productivity,  but are likely due to
environmental changes as several other species have exhibited similar trends in these years (e.g., cod, ling, etc.).  However, this increase and high recruitment uncertainty in recent years should be taken into account when projecting the stock for annual advice: overly optimistic mean recruitment should be avoided.




