---
output: 
  word_document:
    reference_docx: ../../report_template_v1.5.dotx
tables: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}

## Options about the creation of the pdf file
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, error = FALSE, cache = FALSE, 
                      fig.width = 10, fig.asp=1.1)
knitr::opts_knit$set(root.dir = normalizePath(".."))

# output: 
#   word_document:
#     reference_docx: report_template_v1.5.dotx

```

**SPiCT scenarios for the Greater silver smelt (_Argentina silus_) in 5b and 6a. **

authors: Martin Pastoors, Lise Helen Ofstad, Frodi Skuvedal

date: `r format(Sys.time(), '%d/%m/%Y %H:%M')`

**Introduction**

This working document present a series of different assessments using the surplus production model in continous time (SPiCT; Pedersen2016) available as an R package (https://github.com/DTUAqua/spict).

```{r data_pkg, include=FALSE}

# devtools::install_github("DTUAqua/spict/spict")
library(spict)
library(tidyverse)
library(cowplot)
library(stockassessment)

## Read in the data
dat <- readxl::read_xlsx("D:/GIT/wk_WKGSS/aru.27.5b6a/data/ARU_indices 20200205.xlsx")

## run retro or not 
runretro <- FALSE
```

**Descriptions of scenarios**

Scenarios with long catch time series

* Scenario 1: long time-series of catch (1988-2018), including the Faroe DW survey (2014-2018), Faroe Summer survey (1996-2018), Scottish DW survey (1998-2018) and CPUE Faroer fleet (1995-2018). Priors: default
* Scenario 2: long time-series of catch (1988-2018), including the Faroe DW survey (2014-2018), and CPUE Faroer fleet (1995-2018). Priors: default  

Scenarios with shorter catch time series

* Scenario 3: shorter time-series of catch (1995-2018), including the Faroe DW survey (2014-2018), Faroe Summer survey (1996-2018) and CPUE Faroer fleet (1995-2018). Priors: default
* Scenario 4: shorter time-series of catch (1995-2018), including the Faroe DW survey (2014-2018) and CPUE Faroer fleet (1995-2018). Priors: default

Scenarios similar to the SAM assessment

* Scenario 5: shorter time-series of catch (1995-2018), including the Faroe DW survey (2014-2018), Faroe Summer survey (1996-2018) and CPUE combined fleet (2005-2018). Priors: default
* Scenario 6: shorter time-series of catch (1995-2018), including the Faroe DW survey (2014-2018), Faroe Summer survey, Scottish DW survey and CPUE Faroer fleet (1995-2018). Priors: default. This run uses the same data as the SAM assessment. 

**Main results**

Results for each of the scenarios are presented in four pages: input data, estimated model parameters, stock summaries and model diagnostics. Overall results are compared in trends of (log) B over average biomass and F over average fishing mortality. 

The Spict fits appear to be very sensitive to the length of the catch time series and to the data series used. Scenarios 1 and 2 (using the long catch time series) have the drawbacks that the initial catches are not accompanied by survey or CPUE observations. That gives rise to very low estimates of biomass at the beginning of the time series. 

Scenarios 3 and 4 are based on a shorter catch time series (starting in 1995) and using the Faroer CPUE series in combination with the Faroerse surveys. They generally give a somewhat similar biomass trajectory as the SAM assessment. 

Scenarios 5 and 6 are closest to the SAM assessment because using the combined CPUE information (2005 onwards) rather than the longer CPUE series from the Faroes. In these runs only the Faroese summer survey has information that goes back to the beginning of the catch time series. This results in a more smooth pattern in both biomass and fishing mortality. 

The comparison between the spict scenarios and the SAM assessment is shown in the final plot. Biomass seems to be somewhat comparable between spict scenarios 3-64 and the SAM assessment. The fishing mortality estimated by SAM is generally much more variable than the Spict runs. 

The Spict runs are rather sensitive to the length of the time series and the type of biomass information used. The Spict runs are only used in a qualitative way and do not really contradict the results from the SAM assessment.   

##### page break

**Scenario 1**

  + Total catch, long time-series (1988-2018) 
  + Faroe DW survey (2014-2018)
  + Faroe Summer survey (1996-2018)
  + Scottish DW survey (1998-2018); some missing years
  + CPUE Faroer fleet (1995-2018)
  + priors: default


```{r fit_scenario1}

dat2 <- dplyr::filter(dat, year >= 1988)

## Make the input list
inp1 <- list(timeC = dat2$year,                                      ## Timing of catch
               obsC  = dat2$catch,                                     ## Observed catches
               timeI = list(dat2$year + dat2$FO_DW_month / 12,          ## Timing of FO_DW survey index
                            dat2$year + dat2$FO_SS_month / 12,
                            dat2$year + dat2$SC_DW_month / 12,
                            dat2$year),
               obsI = list(dat2$FO_DW,                                 ## Observed indices
                           dat2$FO_SS,
                           dat2$SC_DW,
                           dat2$CPUE_FO),
               optimiser.control = list(iter.max = 1e5,               ## Optimiser options 
                                        eval.max = 1e5),              ## sometimes help converge
               priors = list(                                         ## List of priors (empty, i.e. default priors)
                                                                      ## see possible priors with
                                                                      ## list.possible.priors()
                 ))
## Check input time series, remove missing and zero observations
inp1 <- suppressWarnings(check.inp(inp1))

## Plot input data
plotspict.data(inp1)

## Fit spict
fit1<- fit.spict(inp1)

```


##### page break

```{r}

## Summary of the fit - in the vignette there is a line-by-line description of that summary
fit1

```

##### page break

```{r}

## If the model converged, it reports convergence as 0
## Continue with plotting and diagnostics only if convergence was reached
converged <- fit1$opt$convergence == 0

if (converged) {
  ## Calculate the One Step Ahead (osa) residuals  
  fit1 <- calc.osa.resid((fit1))
  
  ## Make a plot showing relative F, relative B, Kobe plot catch 
  par(mfrow = c(2,2),  ## 2x2 subplots
      mar = c(4.1, 4.1, 0.5, 0.5)) ## Change default margins for the plots 
  plotspict.catch(fit1)
  plotspict.ffmsy(fit1)
  plotspict.bbmsy(fit1)
  plotspict.fb(fit1)
}

if (converged) {
  plotspict.diagnostic(fit1)
}

## If runretro is TRUE, run and plot the retrospective analysis
if (runretro & converged) {
  fit1 <- retro(fit1)
  plotspict.retro(fit1)
}

```

##### page break

**Scenario 2**

  + Total catch, long time-series (1988-2018) 
  + Faroe DW survey (2014-2018)
  + CPUE Faroer fleet (1995-2018)
  + priors: default
            

```{r fit_scenario2}

dat2 <- dplyr::filter(dat, year >= 1988)

## Make the input list
inp2 <- list(timeC   = dat2$year,                                      ## Timing of catch
               obsC  = dat2$catch,                                     ## Observed catches
               timeI = list(dat2$year + dat2$FO_DW_month / 12,          ## Timing of FO_DW survey index
                            dat2$year ),
               obsI = list(dat2$FO_DW,                                 ## Observed indices
                           dat2$CPUE_FO),
               optimiser.control = list(iter.max = 1e5,               ## Optimiser options 
                                        eval.max = 1e5),              ## sometimes help converge
               priors = list(                                         ## List of priors (empty, i.e. default priors)
                                                                      ## see possible priors with
                                                                      ## list.possible.priors()
                 ))
## Check input time series, remove missing and zero observations
inp2 <- suppressWarnings(check.inp(inp2)) 

## Plot input data
plotspict.data(inp2)


```

##### page break

```{r}

## Fit spict
fit2 <- fit.spict(inp2)

## Summary of the fit - in the vignette there is a line-by-line description of that summary
fit2

```

##### page break

```{r}

## If the model converged, it reports convergence as 0
## Continue with plotting and diagnostics only if convergence was reached
converged <- fit2$opt$convergence == 0

if (converged) {
  ## Calculate the One Step Ahead (osa) residuals  
  fit2 <- calc.osa.resid((fit2))
  
  ## Make a plot showing relative F, relative B, Kobe plot catch 
  par(mfrow = c(2,2),  ## 2x2 subplots
      mar = c(4.1, 4.1, 0.5, 0.5)) ## Change default margins for the plots 
  plotspict.catch(fit2)
  plotspict.ffmsy(fit2)
  plotspict.bbmsy(fit2)
  plotspict.fb(fit2)
}

if (converged) {
  plotspict.diagnostic(fit2)
}

  ## If runretro is TRUE, run and plot the retrospective analysis
  if (runretro & converged) {
    fit <- retro(fit2)
    plotspict.retro(fit2)
  }

```

##### page break



**Scenario 3**

  + Total catch, shorter time-series (1995-2018) 
  + Faroe DW survey (2014-2018)
  + Faroe Summer survey (1996-2018)
  + CPUE Faroer fleet (1995-2018)
  + priors: default
  

```{r fit_scenario3}

dat2 <- dplyr::filter(dat, year >= 1995)

## Make the input list
inp3 <- list(timeC   = dat2$year,                                      ## Timing of catch
               obsC  = dat2$catch,                                     ## Observed catches
               timeI = list(dat2$year + dat2$FO_DW_month / 12,          ## Timing of FO_DW survey index
                            dat2$year + dat2$FO_SS_month / 12,
                            dat2$year),
               obsI = list(dat2$FO_DW,                                 ## Observed indices
                           dat2$FO_SS,
                           dat2$CPUE_FO),
               optimiser.control = list(iter.max = 1e5,               ## Optimiser options 
                                        eval.max = 1e5),              ## sometimes help converge
               priors = list(                                         ## List of priors (empty, i.e. default priors)
                                                                      ## see possible priors with
                                                                      ## list.possible.priors()
                 ))
## Check input time series, remove missing and zero observations
inp3 <- suppressWarnings(check.inp(inp3))

## Plot input data
plotspict.data(inp3)



```

##### page break

```{r}

## Fit spict
fit3 <- fit.spict(inp3)

## Summary of the fit - in the vignette there is a line-by-line description of that summary
fit3

```

##### page break

```{r}

## If the model converged, it reports convergence as 0
## Continue with plotting and diagnostics only if convergence was reached
converged <- fit3$opt$convergence == 0

if (converged) {
  ## Calculate the One Step Ahead (osa) residuals  
  fit3 <- calc.osa.resid((fit3))
  
  ## Make a plot showing relative F, relative B, Kobe plot catch 
  par(mfrow = c(2,2),  ## 2x2 subplots
      mar = c(4.1, 4.1, 0.5, 0.5)) ## Change default margins for the plots 
  
  plotspict.catch(fit3)
  plotspict.ffmsy(fit3)
  plotspict.bbmsy(fit3)
  plotspict.fb(fit3)
}

if (converged) {
  plotspict.diagnostic(fit3)
}

  ## If runretro is TRUE, run and plot the retrospective analysis
  if (runretro & converged) {
    fit <- retro(fit3)
    plotspict.retro(fit3)
  }

```

##### page break

**Scenario 4**

  + Total catch, shorter time-series (1995-2018) 
  + Faroe DW survey (2014-2018)
  + CPUE Faroer fleet (1995-2018)
  + priors: default

```{r fit_scenario4}

dat2 <- dplyr::filter(dat, year >= 1995)

## Make the input list
inp4 <- list(timeC = dat2$year,                                      ## Timing of catch
               obsC  = dat2$catch,                                     ## Observed catches
               timeI = list(dat2$year + dat2$FO_DW_month / 12,          ## Timing of FO_DW survey index
                            dat2$year),
               obsI = list(dat2$FO_DW,                                 ## Observed indices
                           dat2$CPUE_FO),
               optimiser.control = list(iter.max = 1e5,               ## Optimiser options 
                                        eval.max = 1e5),              ## sometimes help converge
               priors = list(                                         ## List of priors (empty, i.e. default priors)
                                                                      ## see possible priors with
                                                                      ## list.possible.priors()
                 ))
## Check input time series, remove missing and zero observations
inp4 <- suppressWarnings(check.inp(inp4))

## Plot input dat2a
plotspict.data(inp4)

```

##### page break

```{r}

## Fit spict
fit4<- fit.spict(inp4)

## Summary of the fit - in the vignette there is a line-by-line description of that summary
fit4

```

##### page break

```{r}
## If the model converged, it reports convergence as 0
## Continue with plotting and diagnostics only if convergence was reached
converged <- fit4$opt$convergence == 0

if (converged) {
  ## Calculate the One Step Ahead (osa) residuals  
  fit4 <- calc.osa.resid((fit4))
  
  ## Make a plot showing relative F, relative B, Kobe plot catch 
  par(mfrow = c(2,2),  ## 2x2 subplots
      mar = c(4.1, 4.1, 0.5, 0.5)) ## Change default margins for the plots 
  
  plotspict.catch(fit4)
  plotspict.ffmsy(fit4)
  plotspict.bbmsy(fit4)
  plotspict.fb(fit4)
}

if (converged) {
  plotspict.diagnostic(fit4)
}

## If runretro is TRUE, run and plot the retrospective analysis
if (runretro & converged) {
  fit <- retro(fit4)
  plotspict.retro(fit4)
}

```

##### page break

**Scenario 5**

  + Total catch, shorter time-series (1995-2018) 
  + Faroe DW survey (2014-2018)
  + Faroe summer survey (1995-2018)
  + CPUE combined fleet (2005-2018)
  + priors: default

```{r fit_scenario5}

dat2 <- dplyr::filter(dat, year >= 1995)

## Make the input list
inp5 <- list(timeC = dat2$year,                                       ## Timing of catch
               obsC  = dat2$catch,                                    ## Observed catches
               timeI = list(dat2$year + dat2$FO_DW_month / 12,        ## Timing of FO_DW survey index
                            dat2$year + dat2$FO_SS_month / 12,          
                            dat2$year),
               obsI = list(dat2$FO_DW,                                ## Observed indices
                           dat2$FO_SS,
                           dat2$CPUE_comb),
               optimiser.control = list(iter.max = 1e5,               ## Optimiser options 
                                        eval.max = 1e5),              ## sometimes help converge
               priors = list(                                         ## List of priors (empty, i.e. default priors)
                                                                      ## see possible priors with
                                                                      ## list.possible.priors()
                 ))
## Check input time series, remove missing and zero observations
inp5 <- suppressWarnings(check.inp(inp5))

## Plot input dat2a
plotspict.data(inp5)

```

##### page break

```{r}

## Fit spict
fit5<- fit.spict(inp5)

## Summary of the fit - in the vignette there is a line-by-line description of that summary
fit5

```

##### page break

```{r}
## If the model converged, it reports convergence as 0
## Continue with plotting and diagnostics only if convergence was reached
converged <- fit4$opt$convergence == 0

if (converged) {
  ## Calculate the One Step Ahead (osa) residuals  
  fit5 <- calc.osa.resid((fit5))
  
  ## Make a plot showing relative F, relative B, Kobe plot catch 
  par(mfrow = c(2,2),  ## 2x2 subplots
      mar = c(4.1, 4.1, 0.5, 0.5)) ## Change default margins for the plots 
  
  plotspict.catch(fit5)
  plotspict.ffmsy(fit5)
  plotspict.bbmsy(fit5)
  plotspict.fb(fit5)
}

if (converged) {
  plotspict.diagnostic(fit5)
}

## If runretro is TRUE, run and plot the retrospective analysis
if (runretro & converged) {
  fit <- retro(fit5)
  plotspict.retro(fit5)
}

```

##### page break

**Scenario 6**

  + Total catch, shorter time-series (1995-2018) 
  + Faroe DW survey (2014-2018)
  + Faroe summer survey (1995-2018)
  + Scottish DW survey (1998-2018)
  + CPUE combined fleet (2005-2018)
  + priors: default

```{r fit_scenario6}

dat2 <- dplyr::filter(dat, year >= 1995)

## Make the input list
inp6 <- list(timeC = dat2$year,                                       ## Timing of catch
               obsC  = dat2$catch,                                    ## Observed catches
               timeI = list(dat2$year + dat2$FO_DW_month / 12,        ## Timing of FO_DW survey index
                            dat2$year + dat2$FO_SS_month / 12,          
                            dat2$year + dat2$SC_DW_month / 12,          
                            dat2$year),
               obsI = list(dat2$FO_DW,                                ## Observed indices
                           dat2$FO_SS,
                           dat2$SC_DW, 
                           dat2$CPUE_comb),
               optimiser.control = list(iter.max = 1e5,               ## Optimiser options 
                                        eval.max = 1e5),              ## sometimes help converge
               priors = list(                                         ## List of priors (empty, i.e. default priors)
                                                                      ## see possible priors with
                                                                      ## list.possible.priors()
                 ))
## Check input time series, remove missing and zero observations
inp6 <- suppressWarnings(check.inp(inp6))

## Plot input dat2a
plotspict.data(inp6)

```

##### page break

```{r}

## Fit spict
fit6<- fit.spict(inp6)

## Summary of the fit - in the vignette there is a line-by-line description of that summary
fit6

```

##### page break

```{r}
## If the model converged, it reports convergence as 0
## Continue with plotting and diagnostics only if convergence was reached
converged <- fit4$opt$convergence == 0

if (converged) {
  ## Calculate the One Step Ahead (osa) residuals  
  fit6 <- calc.osa.resid((fit6))
  
  ## Make a plot showing relative F, relative B, Kobe plot catch 
  par(mfrow = c(2,2),  ## 2x2 subplots
      mar = c(4.1, 4.1, 0.5, 0.5)) ## Change default margins for the plots 
  
  plotspict.catch(fit6)
  plotspict.ffmsy(fit6)
  plotspict.bbmsy(fit6)
  plotspict.fb(fit6)
}

if (converged) {
  plotspict.diagnostic(fit6)
}

## If runretro is TRUE, run and plot the retrospective analysis
if (runretro & converged) {
  fit <- retro(fit6)
  plotspict.retro(fit6)
}

```

##### page break

**Comparison of scenarios**

```{r, fig.asp=0.9}

c1 <- 
  data.frame(get.par('logB', fit1)) %>% 
  rownames_to_column(var="year") %>% 
  mutate(year = as.numeric(year)) %>% 
  filter(year >= 1995, year <= 2018) %>% 
  filter((year%%1) == 0) %>% 
  mutate(var="logB", fit = "fit1") %>% 
  mutate(B_Bmsy = est / as.numeric(get.par('logBmsy', fit1)[,"est"])) %>% 
  mutate(B_Bavg = est / mean(est, na.rm=TRUE)) 
  
c2 <- 
  data.frame(get.par('logB', fit2)) %>% 
  rownames_to_column(var="year") %>% 
  mutate(year = as.numeric(year)) %>% 
  filter(year >= 1995, year <= 2018) %>% 
  filter((year%%1) == 0) %>% 
  mutate(var="logB", fit = "fit2") %>% 
  mutate(B_Bmsy = est / as.numeric(get.par('logBmsy', fit2)[,"est"])) %>% 
  mutate(B_Bavg = est / mean(est, na.rm=TRUE)) 

c3 <- 
  data.frame(get.par('logB', fit3)) %>% 
  rownames_to_column(var="year") %>% 
  mutate(year = as.numeric(year)) %>% 
  filter(year >= 1995, year <= 2018) %>% 
  filter((year%%1) == 0) %>% 
  mutate(var="logB", fit = "fit3") %>% 
  mutate(B_Bmsy = est / as.numeric(get.par('logBmsy', fit3)[,"est"])) %>% 
  mutate(B_Bavg = est / mean(est, na.rm=TRUE)) 

c4 <- 
  data.frame(get.par('logB', fit4)) %>% 
  rownames_to_column(var="year") %>% 
  mutate(year = as.numeric(year)) %>% 
  filter(year >= 1995, year <= 2018) %>% 
  filter((year%%1) == 0) %>% 
  mutate(var="logB", fit = "fit4") %>% 
  mutate(B_Bmsy = est / as.numeric(get.par('logBmsy', fit4)[,"est"])) %>% 
  mutate(B_Bavg = est / mean(est, na.rm=TRUE)) 

c5 <-
  data.frame(get.par('logB', fit5)) %>% 
  rownames_to_column(var="year") %>% 
  mutate(year = as.numeric(year)) %>% 
  filter(year >= 1995, year <= 2018) %>% 
  filter((year%%1) == 0) %>% 
  mutate(var="logB", fit = "fit5") %>% 
  mutate(B_Bmsy = est / as.numeric(get.par('logBmsy', fit5)[,"est"])) %>% 
  mutate(B_Bavg = est / mean(est, na.rm=TRUE)) 

c6 <-
  data.frame(get.par('logB', fit6)) %>% 
  rownames_to_column(var="year") %>% 
  mutate(year = as.numeric(year)) %>% 
  filter(year >= 1995, year <= 2018) %>% 
  filter((year%%1) == 0) %>% 
  mutate(var="logB", fit = "fit6") %>% 
  mutate(B_Bmsy = est / as.numeric(get.par('logBmsy', fit6)[,"est"])) %>% 
  mutate(B_Bavg = est / mean(est, na.rm=TRUE)) 

cs <-
  data.frame(stockassessment::tsbtable(stockassessment::fitfromweb("GLL_2019_antest7"))) %>% 
  rownames_to_column(var="year") %>% 
  mutate(year = as.numeric(year)) %>% 
  filter(year >= 1995) %>% 
  mutate(var="logB", fit = "sam") %>% 
  mutate(B_Bavg = log(Estimate) / mean(log(Estimate), na.rm=TRUE)) 


p1 <-
  bind_rows(c1, c2, c3, c4, c5, c6, cs) %>% 
  # View()
  ggplot(aes(x=year,y=B_Bavg)) +
  theme_bw() +
  geom_hline(yintercept=1) +
  geom_line(aes(colour=fit)) +
  labs(title="B_Bavg")

# Fishing mortality

f1 <- 
  data.frame(get.par('logF', fit1)) %>% 
  rownames_to_column(var="year") %>% 
  mutate(year = as.numeric(year)) %>% 
  filter(year >= 1995, year <= 2018) %>% 
  filter((year%%1) == 0) %>% 
  mutate(var="F", fit = "fit1") %>% 
  mutate(est = exp(est)) %>% 
  mutate(F_Fmsy = est / exp(as.numeric(get.par('logFmsy', fit1)[,"est"]))) %>% 
  mutate(F_Favg = est / mean(est, na.rm=TRUE)) 

# plotspict.f(fit3)


f2 <- 
  data.frame(get.par('logF', fit2)) %>% 
  rownames_to_column(var="year") %>% 
  mutate(year = as.numeric(year)) %>% 
  filter(year >= 1995, year <= 2018) %>% 
  filter((year%%1) == 0) %>% 
  mutate(var="F", fit = "fit2") %>% 
  mutate(est = exp(est)) %>% 
  mutate(F_Fmsy = est / exp(as.numeric(get.par('logFmsy', fit2)[,"est"]))) %>% 
  mutate(F_Favg = est / mean(est, na.rm=TRUE)) 

f3 <- 
  data.frame(get.par('logF', fit3)) %>% 
  rownames_to_column(var="year") %>% 
  mutate(year = as.numeric(year)) %>% 
  filter(year >= 1995, year <= 2018) %>% 
  filter((year%%1) == 0) %>% 
  mutate(var="F", fit = "fit3") %>% 
  mutate(est = exp(est)) %>% 
  mutate(F_Fmsy = est / exp(as.numeric(get.par('logFmsy', fit3)[,"est"]))) %>% 
  mutate(F_Favg = est / mean(est, na.rm=TRUE)) 

f4 <- 
  data.frame(get.par('logF', fit4)) %>% 
  rownames_to_column(var="year") %>% 
  mutate(year = as.numeric(year)) %>% 
  filter(year >= 1995, year <= 2018) %>% 
  filter((year%%1) == 0) %>% 
  mutate(var="F", fit = "fit4") %>% 
  mutate(est = exp(est)) %>% 
  mutate(F_Fmsy = est / exp(as.numeric(get.par('logFmsy', fit4)[,"est"]))) %>% 
  mutate(F_Favg = est / mean(est, na.rm=TRUE)) 

f5 <- 
  data.frame(get.par('logF', fit5)) %>% 
  rownames_to_column(var="year") %>% 
  mutate(year = as.numeric(year)) %>% 
  filter(year >= 1995, year <= 2018) %>% 
  filter((year%%1) == 0) %>% 
  mutate(var="F", fit = "fit5") %>% 
  mutate(est = exp(est)) %>% 
  mutate(F_Fmsy = est / exp(as.numeric(get.par('logFmsy', fit5)[,"est"]))) %>% 
  mutate(F_Favg = est / mean(est, na.rm=TRUE)) 

f6 <- 
  data.frame(get.par('logF', fit6)) %>% 
  rownames_to_column(var="year") %>% 
  mutate(year = as.numeric(year)) %>% 
  filter(year >= 1995, year <= 2018) %>% 
  filter((year%%1) == 0) %>% 
  mutate(var="F", fit = "fit6") %>% 
  mutate(est = exp(est)) %>% 
  mutate(F_Fmsy = est / exp(as.numeric(get.par('logFmsy', fit6)[,"est"]))) %>% 
  mutate(F_Favg = est / mean(est, na.rm=TRUE)) 

fs <-
  data.frame(stockassessment::fbartable(stockassessment::fitfromweb("GLL_2019_antest7"))) %>% 
  rownames_to_column(var="year") %>% 
  mutate(year = as.numeric(year)) %>% 
  filter(year >= 1995, year <= 2018) %>% 
  mutate(var="F", fit = "sam") %>% 
  mutate(F_Favg = Estimate / mean(Estimate, na.rm=TRUE)) 

p2 <- 
  bind_rows(f1, f2, f3, f4, f5, f6, fs) %>% 
  # View()
  ggplot(aes(x=year,y=F_Favg)) +
  theme_bw() +
  geom_hline(yintercept=1) +
  geom_line(aes(colour=fit)) +
  labs(title="F_Favg")

cowplot::plot_grid(p1, p2, ncol=1)



```

