---
title: |
  | SPiCT scenarios for the Greater silver smelt (_Argentina silus_)
  | in Subareas 1, 2, and 4, and Division 3.a (Norwegian Sea, North Sea and Skagerrak)
author: "Alexandros Kokkalis, Elvar H. Hallfredsson and Lise Heggebakken"
date: "05/02/2020"
output:
  word_document: default
  pdf_document: default
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
tables: yes
bibliography: bib.bib
---

```{r setup, include=FALSE}
## Options about the creation of the pdf file
knitr::opts_chunk$set(
	error = FALSE,
	fig.height = 5,
	fig.width = 6,
	message = FALSE,
	warning = FALSE,
	cache = TRUE
)
```

## Introduction

This working document present a series of different assessments for greater silver smelt in ICES areas 1,2, 3a and 4, using the surplus production model in continous time (SPiCT; @Pedersen2016) available as an R package (https://github.com/DTUAqua/spict). 

## Data

Input data are total catches since 1988 and biomass indices from tree surveys (depending on scenario); 
  - Acoustic biomass estimates from the Norwegian survey in winter/spring at the eastern slope of the Norwetian Sea. The survey was first concucted in 2009, then 2012, and biennially since then.
  - Acoustic biomass estimates from surveys at the eastern slope of he Norwegian Sea in 1991-1993 publiched by Monstad and Johannessen (2003)
  - Swept area biomass estimates from the Norwegian shrimp survey in the North Sea/Skagerrak. The survey was in October 1984-2002, 2004-2005 in May, 2006-2008 in February, and in January sicne then. Time of year apparantly affects the biomass estimates, and the index was thus split in the analysis. Year 2003 and 2016 are missing.

Biomass is in all cases in tonnes.

## Read in the data

```{r data_pkg}
library(spict)

## Read in the data
dat <- readxl::read_xlsx("../data/GSS_indices270120_AK.xlsx")
#plot(dat$year,dat$catchTOT,type = "l",ylim = c(0,max(na.omit(dat$catchTOT))))
## Sum up the catches from each area to get the total catch
dat$catchTOT <- dat$catch1and2 + dat$catch3 + dat$catch4
#plot(dat$year,dat$catchTOT,type = "l", ylim = c(0,max(na.omit(dat$catchTOT))))
## run retro or not 
runretro <- FALSE
```


## Scenario 1

: Input data for Scenario 1

+-----------------+----------------+------------+----------------------------+
| Input data      | Name           | Range      |  Notes                     |
+=================+================+============+============================+
| Catch           | Total catch    | 1988-2018  |                            |
+-----------------+----------------+------------+----------------------------+
| Biomass indices | Shrimp survey  | 1984--2002 |  Split in two periods      |
|                 |                | 2005--2018 |                            |
+-----------------+----------------+------------+----------------------------+
|                 |                |            |  Default priors            |
+-----------------+----------------+------------+----------------------------+
            

```{r fit_scenario1}
## Choose only the years where the survey was in October 
w <- !is.na(dat$northsea_month) & dat$northsea_month == 10
## Choose only the years where the survey was in January or February 
v <- !is.na(dat$northsea_month) & dat$northsea_month %in% c(1, 2)


## Make the input list
inp_NS <- list(timeC = dat$year,                                      ## Timing of catch
               obsC = dat$catchTOT,                                   ## Observed catches
               timeI = list(dat$year[w] + dat$northsea_month[w] / 12, ## Timing of survey index
                            dat$year[v] + dat$northsea_month[v] / 12),
               obsI = list(dat$northsea_SA[w],                        ## Observed indices
                           dat$northsea_SA[v]),
               optimiser.control = list(iter.max = 1e5,               ## Optimiser options 
                                        eval.max = 1e5),              ## sometimes help converge
               priors = list(                                         ## List of priors (empty, i.e. default priors)
                                                                      ## see possible priors with list.possible.priors()
                 ))
## Check input time series, remove missing and zero observations
inp_NS <- check.inp(inp_NS)

## Plot input data
plotspict.data(inp_NS)

## Fit spict
fit_NS <- fit.spict(inp_NS)

## Summary of the fit - in the vignette there is a line-by-line description of that summary
fit_NS

## If the model converged, it reports convergence as 0
## Continue with plotting and diagnostics only if convergence was reached
converged <- fit_NS$opt$convergence == 0
if (converged) {
  ## Calculate the One Step Ahead (osa) residuals  
  fit_NS <- calc.osa.resid((fit_NS))
  
  ## Make a plot showing relative F, relative B, Kobe plot catch 
  par(mfrow = c(2,2),  ## 2x2 subplots
      mar = c(4.1, 4.1, 0.5, 0.5)) ## Change default margins for the plots 
  plotspict.catch(fit_NS)
  plotspict.ffmsy(fit_NS)
  plotspict.bbmsy(fit_NS)
  plotspict.fb(fit_NS)
}


```

```{r diagnostics_scenario_1, fig.height=8}
if (converged) {
  plotspict.diagnostic(fit_NS)
}
```


```{r retro_scenario_1}
  ## If runretro is TRUE, run and plot the retrospective analysis
  if (runretro & converged) {
    fit_NS <- retro(fit_NS)
    plotspict.retro(fit_NS)
  }
```



## Scenario 2
: Input data for Scenario 2

+-----------------+----------------+------------+----------------------------+
| Input data      | Name           | Range      |  Notes                     |
+=================+================+============+============================+
| Catch           | Total catch    | 1988-2018  |                            |
+-----------------+----------------+------------+----------------------------+
| Biomass indices | Shrimp survey  | 1984--2002 |  Only october period     |
+-----------------+----------------+------------+----------------------------+
| Biomass indices | Acoustic survey  | 2009--2018 |  StoX     |
+-----------------+----------------+------------+----------------------------+
|                 |                |            |  Priors:logn=c(log(2),.5,1), |
+-----------------+----------------+------------+----------------------------+
|                 |                |            |  logbkfrac=c(log(.5),1,1)   |
+-----------------+----------------+------------+----------------------------+

```{r fit_scenario2}
## Choose only the years where the survey was in October 
w <- !is.na(dat$northsea_month) & dat$northsea_month == 10
## Choose only the years where the survey was in January or February 
##v <- !is.na(dat$northsea_month) & dat$northsea_month %in% c(1, 2)      


## Make the input list
inp_NS <- list(timeC = dat$year,                                      ## Timing of catch
               obsC = dat$catchTOT,                                   ## Observed catches
               timeI = list(dat$year[w] + dat$northsea_month[w] / 12, ## Timing of survey index
                            dat$year+3.5/12),
               obsI = list(dat$northsea_SA[w],                        ## Observed indices
                          dat$norwegian_sea_AC_stox),
               optimiser.control = list(iter.max = 1e5,               ## Optimiser options 
                                        eval.max = 1e5),              ## sometimes help converge
               priors = list(                                         ## List of priors (empty, i.e. default priors)
 logn=c(log(2),.5,1),
 logbkfrac=c(log(.5),1,1)
## see possible priors with list.possible.priors()
                 ))
## Check input time series, remove missing and zero observations
inp_NS <- check.inp(inp_NS)

## Plot input data
plotspict.data(inp_NS)

## Fit spict
fit_NS <- fit.spict(inp_NS)

## Summary of the fit - in the vignette there is a line-by-line description of that summary
fit_NS

## If the model converged, it reports convergence as 0
## Continue with plotting and diagnostics only if convergence was reached
converged <- fit_NS$opt$convergence == 0
if (converged) {
  ## Calculate the One Step Ahead (osa) residuals  
  fit_NS <- calc.osa.resid((fit_NS))
  
  ## Make a plot showing relative F, relative B, Kobe plot catch 
  par(mfrow = c(2,2),  ## 2x2 subplots
      mar = c(4.1, 4.1, 0.5, 0.5)) ## Change default margins for the plots 
  plotspict.catch(fit_NS)
  plotspict.ffmsy(fit_NS)
  plotspict.bbmsy(fit_NS)
  plotspict.fb(fit_NS)
}


```

```{r diagnostics_scenario_2, fig.height=8}
if (converged) {
  plotspict.diagnostic(fit_NS)
  plotspict.production(fit_NS)
}
```


```{r retro_scenario_2}
  ## If runretro is TRUE, run and plot the retrospective analysis
  if (runretro & converged) {
    fit_NS <- retro(fit_NS)
    plotspict.retro(fit_NS)
  }
```



## Scenario 3

: Input data for Scenario 3

+-----------------+----------------+------------+----------------------------+
| Input data      | Name           | Range      |  Notes                     |
+=================+================+============+============================+
| Catch           | Total catch    | 1988-2018  |                            |
+-----------------+----------------+------------+----------------------------+
| Biomass indices | Acoustic survey  | 2009--2018 |  StoX     |
+-----------------+----------------+------------+----------------------------+
| Biomass indices | Acoustic survey  | 1990-1993 |  Monstad     |
+-----------------+----------------+------------+----------------------------+
|                 |                |            |  Default priors            |
+-----------------+----------------+------------+----------------------------+

```{r fit_scenario3}
## Choose only the years where the survey was in October 
w <- !is.na(dat$northsea_month) & dat$northsea_month == 10
w[1:4]<-FALSE
## Choose only the years where the survey was in January or February 
##v <- !is.na(dat$northsea_month) & dat$northsea_month %in% c(1, 2)


## Make the input list
inp_NS <- list(timeC = dat$year,                                      ## Timing of catch
               obsC = dat$catchTOT,                                   ## Observed catches
               timeI = list(dat$year+3.5/12, dat$year+3/12),
               obsI = list(dat$norwegian_sea_AC_stox,dat$Norwegian_seaAC_Monstad),
               optimiser.control = list(iter.max = 1e3,               ## Optimiser options 
                                        eval.max = 1e3),              ## sometimes help converge
               priors = list(                                         ## List of priors (empty, i.e. default priors)
 logn=c(log(2),.5,1)
 #logbkfrac=c(log(.5),1,1)
## see possible priors with list.possible.priors()
                 ))
## Check input time series, remove missing and zero observations
inp_NS <- check.inp(inp_NS)

## Plot input data
plotspict.data(inp_NS)

## Fit spict
fit_AC <- fit.spict(inp_NS)

## Summary of the fit - in the vignette there is a line-by-line description of that summary
fit_AC

## If the model converged, it reports convergence as 0
## Continue with plotting and diagnostics only if convergence was reached
converged <- fit_AC$opt$convergence == 0
if (converged) {
  ## Calculate the One Step Ahead (osa) residuals  
  fit_AC <- calc.osa.resid((fit_AC))
  
  ## Make a plot showing relative F, relative B, Kobe plot catch 
  par(mfrow = c(2,2),  ## 2x2 subplots
      mar = c(4.1, 4.1, 0.5, 0.5)) ## Change default margins for the plots 
  plotspict.catch(fit_AC)
  plotspict.ffmsy(fit_AC)
  plotspict.bbmsy(fit_AC)
  plotspict.fb(fit_AC)
}


```

```{r diagnostics_scenario_3, fig.height=8}
if (converged) {
  plotspict.diagnostic(fit_AC)
  plotspict.production(fit_AC)
}
```


```{r retro_scenario_3}
  ## If runretro is TRUE, run and plot the retrospective analysis
  ## runretro=TRUE
if (runretro & converged) {
    fit_AC <- retro(fit_AC)
    plotspict.retro(fit_AC)
  }
```

##Scenario 4

: Input data for Scenario 4

+-----------------+----------------+------------+----------------------------+
| Input data      | Name           | Range      |  Notes                     |
+=================+================+============+============================+
| Catch           | Total catch    | 1988-2018  |                            |
+-----------------+----------------+------------+----------------------------+
| Biomass indices | Shrimp survey  | 1988--2002 |  Only october period     |
+-----------------+----------------+------------+----------------------------+
| Biomass indices | Acoustic survey  | 2009--2018 |  StoX     |
+-----------------+----------------+------------+----------------------------+
| Biomass indices | Acoustic survey  | 1990-1993 |  Monstad     |
+-----------------+----------------+------------+----------------------------+
|                 |                |            |  Default priors            |
+-----------------+----------------+------------+----------------------------+

```{r fit_scenario4}
## Choose only the years where the schrimp survey was in October 
w <- !is.na(dat$northsea_month) & dat$northsea_month == 10
w[1:4]<-FALSE #remove years before 1988 in schrip survey
## Choose only the years where the survey was in January or February 
v <- !is.na(dat$northsea_month) & dat$northsea_month %in% c(1)#c(1, 2)
## Make the input list
inp_NS <- list(timeC = dat$year,                                      ## Timing of catch
               obsC = dat$catchTOT,                                   ## Observed catches dat$year[v] + dat$northsea_month[v] / 12
               timeI = list(dat$year[w] + dat$northsea_month[w] / 12,dat$year+3.5/12, dat$year+3/12),
               obsI = list(dat$northsea_SA[w],dat$norwegian_sea_AC_stox,dat$Norwegian_seaAC_Monstad),
               optimiser.control = list(iter.max = 1e3,               ## Optimiser options 
                                        eval.max = 1e3),              ## sometimes help converge
               
               priors = list(                                         ## List of priors (empty, i.e. default priors)
 logn=c(log(2),.5,1)
 #logbkfrac=c(log(.5),1,1)
## see possible priors with list.possible.priors()
                 ))
## Check input time series, remove missing and zero observations
inp_NS <- check.inp(inp_NS)

## Plot input data
plotspict.data(inp_NS)

## Fit spict
fit_NS <- fit.spict(inp_NS)

## Summary of the fit - in the vignette there is a line-by-line description of that summary
fit_NS

## If the model converged, it reports convergence as 0
## Continue with plotting and diagnostics only if convergence was reached
converged <- fit_NS$opt$convergence == 0
if (converged) {
  ## Calculate the One Step Ahead (osa) residuals  
  fit_NS <- calc.osa.resid((fit_NS))
  
  ## Make a plot showing relative F, relative B, Kobe plot catch 
  par(mfrow = c(2,2),  ## 2x2 subplots
      mar = c(4.1, 4.1, 0.5, 0.5)) ## Change default margins for the plots 
  plotspict.catch(fit_NS)
  plotspict.ffmsy(fit_NS)
  plotspict.bbmsy(fit_NS)
  plotspict.fb(fit_NS)
}


```

```{r diagnostics_scenario_4, fig.height=8}
if (converged) {
  plotspict.diagnostic(fit_NS)
  plotspict.production(fit_NS)
}
```


```{r retro_scenario_4}
  ## If runretro is TRUE, run and plot the retrospective analysis
  #runretro=TRUE
if (runretro & converged) {
    fit_NS <- retro(fit_NS)
    plotspict.retro(fit_NS)
  }
```

### Biomass comparison scenario 3 and 4

```{r fig.cap="Figure XXXX. SPiCT biomass estimates from scenario 3 and 4. NC = SPiCT run with only acoustic indices, AC+SA_Oct = SPiCT run with acoustic incieces and swept area index from North Sea/Skagerrak survy in years when it was conducted in October, AC-StoX = acoustic biomass indx from the Norwegian Sea slope survey."}
bAC <- get.par(parname = "logBBmsy",fit_AC, exp = TRUE)
time <- as.numeric(rownames(bAC))
bNS <- get.par(parname = "logBBmsy",fit_NS, exp = TRUE)
BmsyAC<- get.par(parname = "Bmsy",fit_AC)


meanAC <- mean(bAC[,2][time > 2005 & time < 2019])
meanNS <- mean(bNS[,2][time > 2005 & time < 2019])
AC_StoX<-data.frame(na.omit(cbind(dat$year,dat$norwegian_sea_AC_stox / mean(dat$norwegian_sea_AC_stox ,na.rm = TRUE))))
ylim <- c(0.6, 1.4)
par(mfrow =c(1,1))
plot(time, bNS[,2] / meanNS,type="l", ylim = ylim, xlim=c(2005,2018),ylab = "Bioamass", xlab = "Year")
lines(time, bAC[,2] / meanAC,col="red")
#points(dat$year, dat$norwegian_sea_AC_stox / mean(dat$norwegian_sea_AC_stox, na.rm = TRUE, col="blue"),pch =3 )
lines(AC_StoX, type = "b", col="red",pch=16,lty=2)
legend("bottomright",c("AC+SA_Oct", "AC","AC-StoX"), col = c(1,2,2), seg.len = 4, lty = c(1,1,2),pch = c(NA,NA,16),box.lty=0)


#plot(time, bAC[,2] / meanAC,xlim=c(2005,2018),type = "l",ylim =c(0.6,1.4), ylab = "Biomass AC")

idx2 <- c(which(floor(time) == 2017)[1], which(floor(time) == 2018)[1])
#time[idx2]
idx3 <- c(which(floor(time) == 2014)[1], which(floor(time) == 2015)[1], which(floor(time) == 2016)[1])
#time[idx3]
ratio2_3AC <- mean(bAC[idx2,2]) / mean(bAC[idx3,2])
ratio2_3AC
ratio2_3ACexp <- mean(BmsyAC[2]*exp(bAC[idx2,2])) / mean(BmsyAC[2]*exp(bAC[idx3,2]))
ratio2_3ACexp

```

## Discussion


## Referneces

Monstad, T., & Johannessen, A. (2003). Acoustic recordings of greater silver smelt (_Argentina silus_) in Norwegian waters and west of the British Isles, 1989-94. J. Northw. Atl. Fish. Sci., 31, 339-351.
